---
output: pdf_document
---
 KIPP Chicago \\\\ HSR-Style Student Attrition Report
========================================================
```{r options, echo=FALSE, message=FALSE}

#knit2html("Weekly_Attendance_Report_131106.Rmd", output="test.html", options=c('fragment_only', 'skip_style')); system('cp ./test.html ~/Sites/test.local/'); system('cp -r ./figure/* ~/Sites/test.local/figure/'); system('cd ~/Sites/test.local/ && python updateWarHTML.py')


require(knitr)
options(replace.assign=TRUE,
       width=90
      )
opts_knit$set(
   self.contained=FALSE
  )
opts_chunk$set( 
               #fig.align='center', 
               #dev='tikz', 
               #fig.width=18, 
               #fig.height=10.5, 
               #fig.show='hold', 
               cache=TRUE,
               par=TRUE,
               echo=FALSE,
               message=FALSE,
               warning=FALSE,
               autodepend=FALSE
               )
               
```


```{r load_libraries}
library(ProjectTemplate)
setwd('../')
load.project()

#source("~/Dropbox/Consulting/KIPP Ascend/Data Analysis/MAP/Code/R/MAP_helper_functions.R")
```
```{r calculate_date, cach=FALSE}
date.first  <- "2013-08-19" # first day of school year

date.second <- floor_date(today(), unit="week") 

date.title <- date.second -weeks(1) + days(1) 

title.date<-paste(lubridate::month(date.title, 
                                   label=TRUE, 
                                   abbr=FALSE),
                  " ",
                  lubridate::day(date.title), ", ", 
                  lubridate::year(date.title), sep=""
                  )
```
### Data current through week of `r I(title.date)`


### Transfers 
#### Healthy Schools & Regions transfer totals
Cumulative transfers of students enrolled on October 1st of each year and transferred prior to September 20th of the next year.  Students who enrolled **after October 1st** and transferred during the school year are **not included** in in these tables and figures. 

***Cumulative Transfers***. _Darker colors_ indicate the cumulative number of transfers for reasons other than moving. _Lighter colors_ indicate the additional transfers for students who moved outside of the city of Chicago.  The _angled orange lines_ indicate the rate at which a 10 percent attrition rate accumulates; it is an indicator if we are on track (below the line) or off track (above the line) in a given month.

```{r fig_cumulative_transferes, fig.width=10, fig.height=6}

source('../src/transfer_tables.R')


detach("package:plyr")


enrolled<-rbind(group_by(Enrolled.121003, SCHOOLID) %>% 
     summarise(N=n()) %>% 
        mutate(Year="SY12-13"), 
      group_by(Enrolled.131001, SCHOOLID) %>% 
        summarise(N=n()) %>% 
        mutate(Year="SY13-14")
      )

enrolled$School<-mapply(function(x){ switch(as.character(x),
                           "7810" = "KAMS",
                           "78102" = "KAP",
                           "400146"= "KCCP",
                           "400163" = "KBCP"
                           )
                    },
       enrolled$SCHOOLID
       )

enrolled<-mutate(enrolled, 
                 School=factor(School, 
                               levels=c("KAP", "KAMS", "KCCP", "KBCP")
                               )
                 )

xferplot<-left_join(xferplot, 
                    select(enrolled, -SCHOOLID), 
                    by=c("School", "Year")
                    ) 
  
xferplot <- mutate(xferplot, Pct=round(Value/N*100))

todays_month<-as.character(lubridate::month(today(), label = TRUE, abbr = TRUE))

#remove cumulative transfers passed this month
xferplot2<-xferplot[!(xferplot$Year=="SY13-14" & 
                        xferplot$Month >= todays_month & 
                        xferplot$Variable=="Cumulative Transfers"),]

xferplot2.nm<-xferplot.nm[!(xferplot.nm$Year=="SY13-14" & 
                        xferplot.nm$Month >= todays_month & 
                        xferplot.nm$Variable=="Cumulative Transfers"),]

  ggplot(data=subset(xferplot2, Variable=="Ceiling"), 
       aes(x=Month, y=Value)) + 
  geom_area(data=subset(xferplot2, Variable!="Ceiling"), 
           aes(x=Month, y=CumVal, fill=School, group=School), 
           stat='identity',
           #fill="#439539", 
           width=.5, 
           alpha=.4) + 
  geom_area(data=subset(xferplot2.nm, Variable!="Ceiling"), 
           aes(x=Month, y=CumVal, fill=School, group=School), 
           stat='identity',
           #fill="#439539", 
           width=.5, 
           alpha=1) + 
  geom_line(aes(group=Variable), color="#E27425") + 
  geom_text(data=subset(xferplot2, Variable!="Ceiling"), 
            aes(x=Month, 
                y=Value, 
                group=Variable, 
                label=paste0(Value,"\n(",Pct,"%)")), 
            size=2.5,
            vjust=0) +
    geom_text(data=subset(xferplot2.nm, Variable!="Ceiling"), 
            aes(x=Month, 
                y=Value, 
                group=Variable, 
                label=Value), 
            size=2.5,
            vjust=1) +
  facet_grid(Year~School, scale="free_y") +
  scale_fill_manual(values = c("purple",  #KCCP 
                               "#439539", #KAMS
                               "#60A2D7", #KCCP
                               "#C49A6C"  #KBCP
  )
  ) + theme_bw() + 
  theme(axis.title = element_blank(),
        axis.text.x = element_text(size=7,angle=45, hjust=1),
        axis.text.y = element_text(size=7),
        strip.text = element_text(size=7),
        legend.position="none"
  )     

```

$\pagebreak$

### Count of transfers by type, SY 2013-2014
```{r tbl_transfers_by_type, results='asis'}
kable(xtable(Xfer.table), output=TRUE,  row.names=FALSE)
```

### Percent of transfers, movers versus non-movers, SY 2012-2013

```{r tbl_transfers_hsr, results='asis', echo=FALSE}
HSR.attrtions<-data.frame('School'=c("KAP", "KAMS", "KCCP", "Region", "Network"),
                          'Movers'=c("9%", "5%", "10%", "8%", "2%"),
                          'Non Movers'=c("5%", "10%", "5%","7%", "9%"),
                          'Total' = c("14%", "15%", "15%", "15%", "11%")
                          )


kable(xtable(HSR.attrtions),  output=TRUE,  row.names=FALSE)
```



