 KIPP Chicago Regional Attendance Report
========================================================
```{r options, echo=FALSE}

#knit2html("Weekly_Attendance_Report_131106.Rmd", output="test.html", options=c('fragment_only', 'skip_style')); system('cp ./test.html ~/Sites/test.local/'); system('cp -r ./figure/* ~/Sites/test.local/figure/'); system('cd ~/Sites/test.local/ && python updateWarHTML.py')


require(knitr)
options(replace.assign=TRUE,
       width=60
      )
opts_knit$set(
   self.contained=FALSE
  )
opts_chunk$set( 
               #fig.align='center', 
               #dev='tikz', 
               #fig.width=8, 
               #fig.height=10.5, 
               #fig.show='hold', 
               cache=TRUE,
               par=TRUE,
               echo=FALSE,
               message=FALSE,
               warning=FALSE,
               autodepend=FALSE
               )
               
```


```{r load_libraries}
library(ProjectTemplate)
setwd('../')
load.project()

#source("~/Dropbox/Consulting/KIPP Ascend/Data Analysis/MAP/Code/R/MAP_helper_functions.R")
```
```{r calculate_date, cach=FALSE}
date.first  <- "2013-08-19" # first day of school year

date.second <- floor_date(today(), unit="week") 

date.title <- date.second -weeks(1) + days(1) 

title.date<-paste(lubridate::month(date.title, 
                                   label=TRUE, 
                                   abbr=FALSE),
                  " ",
                  lubridate::day(date.title), ", ", 
                  lubridate::year(date.title), sep=""
                  )
```
Week of `r I(title.date)`
--------------------------------------------------------
### Attendance

```{r Attendance_Summaries}

# Create analytical attendence table
source('../src/attendence_tables.R')


AttRateByWeekBySchool.xtable<-xtable(AttRateByWeekBySchool.table)

# Get last x weeks
if(length(unique(DailyEnrollAttend$WeekOfShortDateLabel)) > 6){
  lastxweeks <- 6
  } else {lastxweeks <-length(unique(DailyEnrollAttend$WeekOfShortDateLabel))
  }
```
#### CPS Impact vs Powerschool YTD Attendance
 
```{r Impact_vs_PowerSchool_Att, results='asis'}

#Get impact attendence from google spreadsheet
googurl<-getURL('https://docs.google.com/spreadsheet/pub?key=0AhdL932iuhIIdHk3SFFSc2ZvUE9MZ1pKNldjZ0JLbXc&single=true&gid=6&range=A1%3AC7&output=csv')

impact<-read.csv(textConnection(googurl))

# Build powerschool number from Attendance 

ps.schools<- Attendance[,list(PowerSchool=round((1-(sum(Absent)/sum(Enrolled)))*100,1)), by=list(School)]
ps.pm<-Attendance[School=="KAMS"|School=="KAP",list(School="KAP/KAMS",PowerSchool=round((1-(sum(Absent)/sum(Enrolled)))*100,1))]
ps.region<-Attendance[,list(School="Region",PowerSchool=round((1-(sum(Absent)/sum(Enrolled)))*100,1))]

ps.all<-rbind(ps.schools, ps.pm, ps.region)

impact<-cbind(ps.all, round(impact[2],1))

impact$Difference<-impact$Impact - impact$PowerSchool

kable(impact, format='html', table.attr='class="table table-responsive table-hover table-condensed table-striped"', output=TRUE)
```



#### Daily Enrollment and Attendance by School (last `r I(lastxweeks)` weeks)
The light gray lines shows total enrollment. The green line demarcates 96% of enrollment (i.e., our regional daily attendance goal).  The black line with dots represents actual attendance, where the dot marks the given weekdays attendance.  
```{r fig_Daily_Attendance, fig.height=6, fig.width = 12,  dev='svg'}
#out.height=225, out.width=525}

last6weeks<-ymd(as.character(floor_date(today() -weeks(lastxweeks), unit="week")+1))

DailyEnrollAttend.plotdata<-subset(DailyEnrollAttend, Date>=last6weeks)
DailyEnrollAttend.plotdata$Day <- wday(DailyEnrollAttend.plotdata$Date)
DailyEnrollAttend.plotdata$Enrolled96Pct <- DailyEnrollAttend.plotdata$Enrolled*.96
DailyEnrollAttend.plotdata.melt<-melt(DailyEnrollAttend.plotdata, id=c("Date", "Day", "School", "WeekOfShortDateLabel"), measure.vars=c("Enrolled", "Enrolled96Pct", "Present"))

DailyEnrollAttend.plotdata.melt$variable<-factor(DailyEnrollAttend.plotdata.melt$variable, labels=c("Enrolled", "96% of Enrolled", "Attended"))


ggplot(DailyEnrollAttend.plotdata.melt, aes(x=Day, y=value)) + 
    geom_step(direction="hv", aes(color=variable)) + 
    geom_point(data=subset(DailyEnrollAttend.plotdata.melt, variable=="Attended"), color="black") +
    scale_x_continuous(breaks = c(2,3,4,5,6), labels=c("M","T","W","R","F")) + #Change numberd week days to lettered
    scale_y_continuous("# of Students") + 
    scale_colour_manual("", values=c("#8D8685", "#439539", "black")) +
    facet_grid(School~WeekOfShortDateLabel, scales="free_y") +
    theme_bw() + 
    theme(legend.position="bottom", 
          strip.text.x=element_text(size=6),
          strip.text.y=element_text(size=6),
          axis.text.x=element_text(size=7),
          axis.text.y=element_text(size=5),
           axis.title.x=element_text(size=8),
          axis.title.y=element_text(size=8),
          legend.text=element_text(size=7)
          )


```

#### Weekly & YTD Attendance by School

```{r tbl_Weekly_Attendance, results='asis'}
AttRateByWeekBySchool.xtable<-xtable(arrange(AttRateByWeekBySchool.table, desc(AttRateByWeekBySchool.table[["Week of"]])))

kable(AttRateByWeekBySchool.xtable, format='html', table.attr='class="table table-responsive table-hover table-condensed table-striped"', output=TRUE)
```


#### Student Attendance 
These tables list the students in each grade who are  among the bottom 10% in terms of **Average Daily Attendance (ADA)**.  ADA is calcualte separately for each student and simply gives the percentage of time a student has attended school *since they enrolled at the school*. Consequenlty, a recently enrolled student missing only 5 days would appear on this list if they've only been enrolled 15 days.  Such a student would have a calculated ADA of 66.6%, meaning they've attended school two thirds of the time since they enrolled and have missed 1/3 of the days since enrolling. 

The set of students in the 10th percentile of lower is calculated separately for each grade level at each school.

##### KAP
```{r Student_Attendance_KAP, results='asis'}
kable(AttByStudentBySchool[School=="KAP",c("Grade", "Student", "ADA", "ADA (prior month)", "Absences"), with=FALSE], format='html', table.attr='class="table table-responsive table-hover table-condensed table-striped"', output=TRUE)
```

##### KAMS
```{r Student_Attendance_KAMS, results='asis'}
kable(AttByStudentBySchool[School=="KAMS",c("Grade", "Student", "ADA", "ADA (prior month)", "Absences"), with=FALSE], format='html', table.attr='class="table table-responsive table-hover table-condensed table-striped"', output=TRUE)
```

##### KCCP
```{r Student_Attendance_KCCP, results='asis'}
kable(AttByStudentBySchool[School=="KCCP",c("Grade", "Student", "ADA", "ADA (prior month)", "Absences"), with=FALSE], format='html', table.attr='class="table table-responsive table-hover table-condensed table-striped"', output=TRUE)
```

##### KBCP
```{r Student_Attendance_KBCP, results='asis'}
kable(AttByStudentBySchool[School=="KBCP",c("Grade", "Student", "ADA", "ADA (prior month)", "Absences"), with=FALSE], format='html', table.attr='class="table table-responsive table-hover table-condensed table-striped"', output=TRUE)
```

### Suspensions
In this report the metric **suspensions** represents days absent due to suspensions and *not* the number of suspension events. That is, a student recieving a 3-day suspension would appear in this report as having 3 "suspensions" (i.e., three absenses due to being out of school serving a suspension).
```{r Suspensions, cache=FALSE, echo=FALSE, results='hide'}

source('../src/suspension_tables.R')

```
#### Total YTD Absenses due to Suspensions 
##### by School 
```{r tbl_Susps_by_School, results='asis'}
kable(YTDSuspensionsByGradeBySchool.xtable, format='html', table.attr='class="table table-responsive table-hover table-condensed table-striped"', output=TRUE,  row.names=FALSE)
```
##### by Week

```{r tbl_Susps_by_week, results='asis'}
kable(WeeklySuspensionsBySchool.xtable, format='html', table.attr='class="table table-responsive table-hover table-condensed table-striped"', output=TRUE,  row.names=FALSE)
```

#### YTD Students Suspended (ordered by number of absenses due to suspensions)
```{r tbl_Suspensions_Leaderboard, results='asis'}

#names(Sups.leaders)<-c("Student", "School", "Grade", "Suspensions")


kable(Sups.leaders,format='html', table.attr='class="table table-responsive table-hover table-condensed table-striped"', output=TRUE,  row.names=FALSE)
```

### Transfers 
#### Healthy Schools & Regions transfer totals
Cumulative transfers of students enrolled on October 1st of each year and transferred prior to September 20th of the next year.  Students who enrolled **after October 1st** and transferred during the school year are **not included** in in these tables and figures. 
```{r fig_cumulative_transferes, fig.height=5, fig.width=10,  dev='svg'}
source('../src/transfer_tables.R')
       
  ggplot(data=subset(xferplot, Variable=="Ceiling"), 
       aes(x=Month, y=Value)) + 
  geom_line(aes(group=Variable), color="#E27425") + 
  geom_bar(data=subset(xferplot, Variable!="Ceiling"), 
           aes(x=Month, y=Value, fill=School), 
           stat='identity',
           #fill="#439539", 
           width=.5) + 
  geom_text(data=subset(xferplot, Variable!="Ceiling"), 
            aes(x=Month, y=Value-.5, group=Variable, label=Value), 
            size=3,
            vjust=1) +
  facet_grid(Year~School, scale="free_y") +
  scale_fill_manual(values = c("purple",  #KCCP 
                               "#439539", #KAMS
                               "#60A2D7", #KCCP
                               "#A7CFEE"  #KBCP
  )
  ) + theme_bw() + 
  theme(axis.title = element_blank(),
        axis.text.x = element_text(size=7,angle=45, hjust=1),
        axis.text.y = element_text(size=7),
        strip.text = element_text(size=7),
        legend.position="none"
  )     

```
```{r tbl_transfers_by_type, results='asis'}
kable(xtable(Xfer.table),format='html', table.attr='class="table table-responsive table-hover table-condensed table-striped"', output=TRUE,  row.names=FALSE)
```

#### Students Transfered by School
```{r tbl_studens_transferred_list, results='asis'}
  kable(xtable(Xfer.students.table),format='html', table.attr='class="table table-responsive table-hover table-condensed table-striped"', output=TRUE,  row.names=FALSE)
```
