KIPP Chicago Regional Attendance Report
========================================================
Week of February 18, 2013
--------------------------------------------------------

### Attendence
```{r options, echo=FALSE}
options(replace.assign=TRUE,width=60)
opts_chunk$set( 
               fig.align='center', 
               #dev='tikz', 
               fig.width=8, 
               fig.height=10.5, 
               fig.show='hold', 
               cache=FALSE, 
               par=TRUE,
               echo=FALSE,
               message=FALSE,
               warning=FALSE,
               autodepend=TRUE
               )
```
```{r load_libraries}
setwd("~/Dropbox/Consulting/KIPP Ascend/Data Analysis/Weekly Attendance Report/")



library(plyr)  #To manipulate data
library(reshape) #More data manipulation
library(ggplot2) #Graphics of grammer graphing
library(grid) #More Graphing
library(gridExtra) #better than par(mfrow=c(r,c)) for arranging ggplot2 and lattice graphics
library(lubridate) #for dealing with dates
library(xtable) #for pretty tables
library(RJDBC) #To get data form PowerSchool DB
library(RODBC) #To get data form PowerSchool DB
#source MAP helper functions
source("~/Dropbox/Consulting/KIPP Ascend/Data Analysis/MAP/Code/R/MAP_helper_functions.R")
```
```{r establsih_DB_connections, cache=FALSE}
drvr <- JDBC("oracle.jdbc.driver.OracleDriver", "/Users/chaid/Dropbox/JDBC Drivers/ojdbc5.jar","")
pscon <- dbConnect(drvr,"jdbc:oracle:thin:@10.160.29.47:1521:IL039","psnavigator","laidephy")
```
```{r get_Attendance_Data, cache=TRUE}
#Get Attendance data.  
# This SQL statement retrieves each enrolled student and their attendence status form PowerSchool 
# by date for a given dateThi sis a table with each student for each day with status for the day
Attendence<-dbGetQuery(pscon, "SELECT 
    m.schoolid, 
  	m.grade_level,
  	m.calendardate, 
	m.STUDENT_NUMBER AS StudentID, 
  	m.lastfirst,
	m.Enrolled,
	a.Att_Code,
	a.Description as AttDescr,
	a.Presence_Status_CD,
	a.COURSE_CREDIT_POINTS,
	CASE 
		WHEN	a.Presence_Status_CD = 'Absent' THEN a.COURSE_CREDIT_POINTS 
		ELSE 0 
	END as Absent
FROM (
	SELECT
		psmd.SchoolID,  
		psmd.grade_level, 
		psmd.calendardate, 
		psmd.studentid,
		s.STUDENT_NUMBER,
		s.LASTFIRST, 
		1 as Enrolled  
	FROM PS_Membership_Defaults psmd
	LEFT JOIN students s ON psmd.StudentID = s.id
	Where 	calendardate >= '13-AUG-12'
		AND  calendardate <= '24-FEB-13'
) m
LEFT JOIN (
	SELECT 
		att.schoolid, 
		att.StudentID,
		att.Att_Date,
		attc.Att_Code,
		attc.Description,
		attc.Presence_Status_CD,
		attc.COURSE_CREDIT_POINTS
	FROM Attendance att
  	INNER JOIN Attendance_Code attc ON att.Attendance_CodeID = attc.ID
	WHERE 
		att.Att_Mode_Code = 'ATT_ModeDaily'
  		AND att.Att_Date >= '13-AUG-12'
  		AND att.Att_Date <= '24-Feb-13'
		AND (attc.att_code = 'A' OR attc.att_code = 'S' or attc.att_code = 'X' or attc.att_code = 'H')
) a
ON m.STUDENTID = a.studentid AND m.calendardate =a.Att_Date AND m.schoolID = a.schoolid
ORDER BY schoolid, grade_level, calendardate
")
```
```{r polish_Attendance, cache=FALSE}

Attendence$SchoolInitials[Attendence$SCHOOLID==7810]<-"KAMS"
Attendence$SchoolInitials[Attendence$SCHOOLID==78102]<-"KAPS"
Attendence$SchoolInitials[Attendence$SCHOOLID==400146]<-"KCCP"
Attendence$SchoolInitials<-factor(Attendence$SchoolInitials, levels=c("KAPS", "KAMS", "KCCP"))


#Summarize enrollments and absences by day



DailyEnrollAttend<-ddply(Attendence, .(SCHOOLID, CALENDARDATE), summarise, Enrolled=sum(ENROLLED), Absent=sum(ABSENT))

DailyEnrollAttendByGrade<-ddply(Attendence, .(SCHOOLID, GRADE_LEVEL, CALENDARDATE), summarise, Enrolled=sum(ENROLLED), Absent=sum(ABSENT))
#rename some columns (becuase ALL CAPS is annoying)

names(DailyEnrollAttend)<-c("SchoolID", "Date", "Enrolled", "Absent")
names(DailyEnrollAttendByGrade)<-c("SchoolID", "Grade","Date", "Enrolled", "Absent")


#transform dates
DailyEnrollAttend$Date<-ymd_hms(DailyEnrollAttend$Date)
DailyEnrollAttendByGrade$Date<-ymd_hms(DailyEnrollAttendByGrade$Date)


#Some quick daily stats
DailyEnrollAttend$Present<-DailyEnrollAttend$Enrolled-DailyEnrollAttend$Absent
DailyEnrollAttendByGrade$Present<-DailyEnrollAttendByGrade$Enrolled-DailyEnrollAttendByGrade$Absent

DailyEnrollAttend$PctAbsent<-DailyEnrollAttend$Absent/DailyEnrollAttend$Enrolled
DailyEnrollAttendByGrade$PctAbsent<-DailyEnrollAttendByGrade$Absent/DailyEnrollAttendByGrade$Enrolled

DailyEnrollAttend$PctPresent<-1-DailyEnrollAttend$PctAbsent
DailyEnrollAttendByGrade$PctPresent<-1-DailyEnrollAttendByGrade$PctAbsent

DailyEnrollAttend$PctPresentGTE95<-DailyEnrollAttend$PctPresent>=.95
DailyEnrollAttendByGrade$PctPresentGTE95<-DailyEnrollAttendByGrade$PctPresent>=.95

DailyEnrollAttend$Present96PctBenchMark<-.96*DailyEnrollAttend$Enrolled
DailyEnrollAttendByGrade$Present96PctBenchMark<-.96*DailyEnrollAttendByGrade$Enrolled

DailyEnrollAttend$WeekInYear<-week(DailyEnrollAttend$Date)
DailyEnrollAttendByGrade$WeekInYear<-week(DailyEnrollAttendByGrade$Date)

DailyEnrollAttend$WeekInSchoolYear<-(intv<-floor_date(DailyEnrollAttend$Date, unit="week")-min(floor_date(DailyEnrollAttend$Date, unit="week")))/dweeks(1)+1
DailyEnrollAttendByGrade$WeekInSchoolYear<-DailyEnrollAttendByGrade$WeekInYear - min(DailyEnrollAttendByGrade$WeekInYear) +1

DailyEnrollAttend$WeekOfDate<-floor_date(DailyEnrollAttend$Date, unit="week") + days(1) 
DailyEnrollAttendByGrade$WeekOfDate<-floor_date(DailyEnrollAttendByGrade$Date, unit="week") + days(1) 

#Long Week (of) label
DailyEnrollAttend$WeekOfDateLabel<-paste("Week of \n", month(DailyEnrollAttend$WeekOfDate,label=TRUE, abbr=TRUE), day(DailyEnrollAttend$WeekOfDate), sep=" ")
DailyEnrollAttendByGrade$WeekOfDateLabel<-paste("Week of \n", month(DailyEnrollAttendByGrade$WeekOfDate,label=TRUE, abbr=TRUE), day(DailyEnrollAttendByGrade$WeekOfDate), sep=" ")

DailyEnrollAttend$WeekOfDateLabel<-factor(DailyEnrollAttend$WeekInSchoolYear, labels=unique(DailyEnrollAttend$WeekOfDateLabel))
DailyEnrollAttendByGrade$WeekOfDateLabel<-factor(DailyEnrollAttendByGrade$WeekInSchoolYear, labels=unique(DailyEnrollAttendByGrade$WeekOfDateLabel))

#Short Week  Label
DailyEnrollAttend$WeekOfShortDateLabel<-paste(month(DailyEnrollAttend$WeekOfDate,label=TRUE, abbr=TRUE), day(DailyEnrollAttend$WeekOfDate), sep=" ")
DailyEnrollAttendByGrade$WeekOfShortDateLabel<-paste(month(DailyEnrollAttendByGrade$WeekOfDate,label=TRUE, abbr=TRUE), day(DailyEnrollAttendByGrade$WeekOfDate), sep=" ")


DailyEnrollAttend$WeekOfShortDateLabel<-factor(DailyEnrollAttend$WeekInSchoolYear, labels=unique(DailyEnrollAttend$WeekOfShortDateLabel))
DailyEnrollAttendByGrade$WeekOfShortDateLabel<-factor(DailyEnrollAttendByGrade$WeekInSchoolYear, labels=unique(DailyEnrollAttendByGrade$WeekOfShortDateLabel))


#add School Initials for graphics
DailyEnrollAttend$SchoolInitials[DailyEnrollAttend$SchoolID==7810]<-"KAMS"
DailyEnrollAttendByGrade$SchoolInitials[DailyEnrollAttendByGrade$SchoolID==7810]<-"KAMS"

DailyEnrollAttend$SchoolInitials[DailyEnrollAttend$SchoolID==78102]<-"KAPS"
DailyEnrollAttendByGrade$SchoolInitials[DailyEnrollAttendByGrade$SchoolID==78102]<-"KAPS"

DailyEnrollAttend$SchoolInitials[DailyEnrollAttend$SchoolID==400146]<-"KCCP"
DailyEnrollAttendByGrade$SchoolInitials[DailyEnrollAttendByGrade$SchoolID==400146]<-"KCCP"

DailyEnrollAttend$SchoolInitials<-factor(DailyEnrollAttend$SchoolInitials, levels=c("KAPS", "KAMS", "KCCP"))
DailyEnrollAttendByGrade$SchoolInitials<-factor(DailyEnrollAttendByGrade$SchoolInitials, levels=c("KAPS", "KAMS", "KCCP"))
#head(DailyEnrollAttend)
```
```{r Attendance_Summaries}
#calculate attendence rate by week by school
AttRateByWeekBySchool<-ddply(DailyEnrollAttend,.(SchoolInitials,WeekOfShortDateLabel), summarise, AttRate=sum(Present)/sum(Enrolled)*100)

AttRateYTDBySchool<-ddply(DailyEnrollAttend,.(SchoolInitials), summarise, AttRate=sum(Present)/sum(Enrolled)*100)

AttRateYTDBySchool$WeekOfShortLabel<-"YTD Each School"

AttRateByWeekBySchool.table<-cast(AttRateByWeekBySchool, WeekOfShortDateLabel ~ SchoolInitials)

AttRateYTDBySchool<-reshape(AttRateYTDBySchool, idvar="WeekOfShortLabel",timevar="SchoolInitials", direction="wide")
names(AttRateYTDBySchool)<-c("WeekOfShortDateLabel", "KAPS", "KAMS", "KCCP")




AttRateYTDRegion<-data.frame(WeekOfShortDateLabel="YTD Region", KAPS=sum(DailyEnrollAttend$Present)/sum(DailyEnrollAttend$Enrolled)*100, KAMS=NA,KCCP=NA)

AttRateYTDKAPSKAMS<-data.frame(WeekOfShortDateLabel="YTD KAPS & KAMS", KAPS=sum(subset(DailyEnrollAttend, SchoolInitials!="KCCP")$Present)/sum(subset(DailyEnrollAttend, SchoolInitials!="KCCP")$Enrolled)*100, KAMS=NA,KCCP=NA)
  
  
AttRateByWeekBySchool.table<-rbind(AttRateByWeekBySchool.table,AttRateYTDBySchool,AttRateYTDKAPSKAMS, AttRateYTDRegion)
AttRateByWeekBySchool.table[,c(2:4)] <- round(AttRateByWeekBySchool.table[,c(2:4)],1)  




names(AttRateByWeekBySchool.table)[1]<-"Week of" #a better column title
AttRateByWeekBySchool.xtable<-xtable(AttRateByWeekBySchool.table)
```

#### Daily Enrollment and Attendance by School (last 20 weeks)
The greeen line demarcates 95% of enrollment (i.e., our regional daily attendance goal).
```{r fig_Daily_Attendence, fig.height=3, fig.width = 7, dpi=300, out.height=450, out.width=1050}

last15weeks<-ymd(as.character(floor_date(today() -weeks(20), unit="week")+1))

DailyEnrollAttend.plotdata<-subset(DailyEnrollAttend, Date>=last15weeks)
DailyEnrollAttend.plotdata$Day <- wday(DailyEnrollAttend.plotdata$Date)
DailyEnrollAttend.plotdata$Enrolled95Pct <- DailyEnrollAttend.plotdata$Enrolled*.95
DailyEnrollAttend.plotdata.melt<-melt(DailyEnrollAttend.plotdata, id=c("Date", "Day", "SchoolInitials", "WeekOfDateLabel"), measure.vars=c("Enrolled", "Enrolled95Pct", "Present"))

DailyEnrollAttend.plotdata.melt$variable<-factor(DailyEnrollAttend.plotdata.melt$variable, labels=c("Enrolled", "95% of Enrolled", "Attended"))


ggplot(DailyEnrollAttend.plotdata.melt, aes(x=Day, y=value)) + 
    geom_step(direction="hv", aes(color=variable)) + 
    scale_x_continuous(breaks = c(2,3,4,5,6), labels=c("M","T","W","R","F")) + #Change numberd week days to lettered
    scale_y_continuous("# of Students") + 
    scale_colour_manual("", values=c("#8D8685", "#439539", "black")) +
    facet_grid(SchoolInitials~WeekOfDateLabel, scales="free_y") +
    theme_bw() + 
    theme(legend.position="bottom", 
          strip.text.x=theme_text(size=4.75),
          axis.text.x=theme_text(size=4))


```

#### Weekly & YTD Attendance by School
```{r tbl_Weekly_Attendance, dpi=300, out.width=400, fig.width=4, fig.height=9, fig.align='left'}
grid.table(AttRateByWeekBySchool.xtable, show.rownames=FALSE)
```

#### Student Attendance 
```{r Student_Attendence, fig.width=11, fig.height=20}
AttByStudentBySchool<-arrange(ddply(Attendence, .(LASTFIRST, SchoolInitials, GRADE_LEVEL), summarise, Absences=sum(ABSENT)), SchoolInitials, desc(Absences))

AttByStudentBySchool<-na.omit(AttByStudentBySchool)
names(AttByStudentBySchool)<-c("Student", "School", "Grade", "Absences")


leaderboards<-function(df, schoolname="KAMS",height=c(.3,.7)){
  x<-df[df$School==schoolname,]
  x<-x[,-2]
  
  t.title<-textGrob(schoolname, just="bottom")
  t.table<-tableGrob(x[1:15,], show.rownames=FALSE, as.table=TRUE)
 
  arrangeGrob(t.title, t.table, heights=height, default.units="npc")
}

grid.arrange(leaderboards(AttByStudentBySchool, schoolname="KAPS", height=c(.1,.5)),leaderboards(AttByStudentBySchool, schoolname="KAMS", height=c(.1,.5)),leaderboards(AttByStudentBySchool, schoolname="KCCP", height=c(.1,.5)), ncol=2, nrow=2, heights=unit(10, "inches"))
```


### Suspensions
```{r Suspensions, cache=FALSE}
Suspensions<-subset(Attendence, ATT_CODE=='S')

Suspensions$WeekInYear<-week(Suspensions$CALENDARDATE)



Suspensions$WeekInSchoolYear<-(floor_date(ymd_hms(Suspensions$CALENDARDATE), unit="week")-min(floor_date(ymd_hms(Suspensions$CALENDARDATE), unit="week")))/dweeks(1)+1

Suspensions$WeekOfDate<-floor_date(ymd_hms(Suspensions$CALENDARDATE), unit="week") + days(1) 

Suspensions<-arrange(Suspensions, CALENDARDATE)

Suspensions$WeekOfShortDateLabel<-paste(month(Suspensions$WeekOfDate,label=TRUE, abbr=TRUE), day(Suspensions$WeekOfDate), sep=" ")
Suspensions$WeekOfShortDateLabel<-factor(Suspensions$WeekInSchoolYear, labels=unique(Suspensions$WeekOfShortDateLabel))

Suspensions$SchoolInitials[Suspensions$SCHOOLID==7810]<-"KAMS"
Suspensions$SchoolInitials[Suspensions$SCHOOLID==78102]<-"KAPS"
Suspensions$SchoolInitials[Suspensions$SCHOOLID==400146]<-"KCCP"
Suspensions$SchoolInitials<-factor(Suspensions$SchoolInitials, levels=c("KAPS", "KAMS", "KCCP"))

DailySuspensionByGradeByWeek<-ddply(Suspensions, .(SchoolInitials, GRADE_LEVEL, WeekOfShortDateLabel), summarise, Suspended=sum(ABSENT))

#weekly suspension by school 
WeeklySuspensionsBySchool.table<-cast(DailySuspensionByGradeByWeek, WeekOfShortDateLabel~SchoolInitials, sum, margins=TRUE)

# Change Week of row (all) to Total
levels(WeeklySuspensionsBySchool.table$WeekOfShortDateLabel)[levels(WeeklySuspensionsBySchool.table$WeekOfShortDateLabel)=="(all)"]<-"Total"

names(WeeklySuspensionsBySchool.table)<-c("Week of", "KAPS", "KAMS", "Total")
levels(WeeklySuspensionsBySchool.table[,1])[levels(WeeklySuspensionsBySchool.table[,1])=="(all)"]<-"Total"

WeeklySuspensionsBySchool.xtable<-xtable(WeeklySuspensionsBySchool.table, digits=0)


#YTD Suspsesnions by Grade By School
YTDSuspensionsByGradeBySchool.table<-cast(DailySuspensionByGradeByWeek, GRADE_LEVEL~SchoolInitials, sum, margins=TRUE)

names(YTDSuspensionsByGradeBySchool.table)<-c("Grade", "KAPS", "KAMS", "Total")
levels(YTDSuspensionsByGradeBySchool.table[,1])[levels(YTDSuspensionsByGradeBySchool.table[,1])=="(all)"]<-"Total"

YTDSuspensionsByGradeBySchool.xtable<-xtable(YTDSuspensionsByGradeBySchool.table, digits=0)



#YTD Suspsesnions by Grade by Week
YTDSuspensionsByWeekByGrade.table<-cast(DailySuspensionByGradeByWeek, WeekOfShortDateLabel~GRADE_LEVEL, sum, margins=TRUE)

names(YTDSuspensionsByWeekByGrade.table)<-c("Week of", "1st", "2nd", "5th","6th", "7th", "8th", "Total")
levels(YTDSuspensionsByWeekByGrade.table[,1])[levels(YTDSuspensionsByWeekByGrade.table[,1])=="(all)"]<-"Total"

YTDSuspensionsByWeekByGrade.xtable<-xtable(YTDSuspensionsByWeekByGrade.table, digits=0)
```
#### Total YTD Suspensions by School & by Week
```{r tbl_Suspensions, fig.height=6}
tbl.susp.1<-tableGrob(YTDSuspensionsByGradeBySchool.xtable, show.rownames=FALSE, as.table=TRUE,main="YTD Suspsensions by School")
tbl.susp.2<-tableGrob(YTDSuspensionsByWeekByGrade.xtable,show.rownames=FALSE)

grid.arrange(tbl.susp.1,tbl.susp.2, ncol=2)
```
#### YTD Students Suspended (ordered by number of suspensions)
```{r tbl_Suspensions_Leaderboard, fig.width=11, fig.height=12}
Sups.leaders<-arrange(ddply(Suspensions, .(LASTFIRST, SchoolInitials, GRADE_LEVEL), summarise, Suspensions=sum(ABSENT)), desc(Suspensions))
names(Sups.leaders)<-c("Student", "School", "Grade", "Suspensions")
grid.table(Sups.leaders)
```
#### Students Suspended, Previous 4 Weeks (ordered by number of suspensions)
```{r tbl_Suspension_most_recent_week, fig.height=6}

last4weeks<-ymd(as.character(floor_date(today() -weeks(4), unit="week")+1))

Sups.lastweek<-arrange(subset(Suspensions, WeekOfDate>=last4weeks, select=c(CALENDARDATE, GRADE_LEVEL, LASTFIRST, SchoolInitials)),SchoolInitials, GRADE_LEVEL, CALENDARDATE, LASTFIRST)
Sups.lastweek$Date<-paste(month(Sups.lastweek$CALENDARDATE, label=TRUE),day(Sups.lastweek$CALENDARDATE), sep=" ") 

Sups.lastweek.table<-Sups.lastweek[,c("Date", "SchoolInitials","GRADE_LEVEL", "LASTFIRST")]
names(Sups.lastweek.table)<-c("Date", "School","Grade", "Student Name")
grid.table(Sups.lastweek.table)

```


