---
title: "Program Evaluation in the Real World Translated into R"
author: "Chris Haid"
date: "June 24, 2015"
output: html_document
---

This document recreates the statistical techniques that Sarah Cohedes presented to the KIPP Data Retreat on June 24, 2015.


# Prereqs

We will first want to load the data sets from CSV files into our `R` session. 
Since the files are csv comment we will use the approprialtely named `read.csv()` function to assign each csv file, which is effectively a table, to a `data.frame` in R:

```{r laod_data}
 practice_FE <- read.csv("practice_FE.CSV")
 practice_reg <- read.csv("practice_reg.csv")
 practice_dd <- read.csv("practice_dd.csv")
 practice_rd <- read.csv("practic_rd.csv")
```
We can take a look at theses data sets by running `summary()` on each of them:

```{r summary_data}
 summary(practice_FE)
 
 require(dplyr)
 glimpse(practice_FE)
```

# Regression Discontinuity Design 
```{r RD}

 require(rdd)
 require(dplyr)
 
 rd_data <- practice_rd %>%
   mutate(assigned_stmath = ifelse(diagnostic <= 40, 1, 0),
          dtt = diagnostic- 40)
 
 
 mod1<-
 
 summary(mod1)
 
 plot(mod1)
 
mod_bw2 <- RDestimate(test ~ dtt, cutpoint = 0, data = rd_data, bw=4)
 
# Rolling our own
bw <- 1
d_l <- rd_data %>% 
  filter(dtt >= -bw*sd(dtt),
         dtt <= 0)

d_r <- rd_data %>% 
  filter(dtt <= bw*sd(dtt),
         dtt >= 0)

 
mod_l <- lm(test ~ dtt, data = d_l)
            
mod_r <- lm(test ~ dtt, data = d_r)

# LATE
coef(mod_l)[[1]] - coef(mod_r)[[1]]


ggplot(d_l, aes(y = test, x = dtt)) +
  stat_smooth(method = "lm") + 
  geom_point() +
  stat_smooth(data = d_r, method="lm") + 
  geom_point(data = d_r) + 
  geom_vline(aes(xintercept = 0))
 
ggplot(rbind_list(d_l, d_r), aes(x = dtt)) + 
  geom_histogram()
 
```

# Difference-in-Difference Analysis

$Y = \alpha + \beta\text{post}\times\text{Treat}
```{r}
glimpse(practice_dd)

practice_dd %>% filter(grade==8) %>% select(grade) 

mod_dd <- lm(test ~ st_math + factor(year) + factor(classid), data = practice_dd)

mod_dd <- lm(test ~ st_math + I(year-min(year))*factor(classid))

mod_dd

```









# Fixed Effects Analysis
```{r fe}
fe_model <- lm(test ~  + cluster(st_math))
```

# Regression Analsyis 


You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
