---
title: "Attendance Analysis and Modelling"
author: "Steph Oliva and Chris Haid"
output:
  html_document:
    df_print: paged
---

This is a first pass at modeling student-level attendence.  The primary use cases are:
* to provide predictions of end-of-year average daily attendance (ADA) by for each student earlier in the year (say the middle of the 1st quarter).
* to provide predictions of end-of-year aveage daily attendance (ADA) by for each school and at the school level earlier in the year (say the middle of the 1st quarter).
* Prepare for the addition of an On-Track metric in SQRP. 

Another goal is gain familiarity with the [tidymodels](https://github.com/tidymodels) suite of packages (e.g, `recipes`, `parsnip`, `resample`)
```{r packages, }
library(tidyverse)
library(silounloadr)
library(kippcolors)
library(janitor)
library(lubridate)
library(caret)
library(broom)
library(modelr)
library(tidymodels)

theme_set(theme_kipp_light())


bigrquery::set_service_token("kipp-chicago-silo-2-aa786970aefd.json")

```
```{r knitr_options}
knitr::opts_knit$set(comment = FALSE, 
                     warning = FALSE, 
                     progress = FALSE, 
                     verbose = FALSE)

knitr::opts_chunk$set(fig.width = 10)
```


Here we grab basic attendances and enrollment data from Silo (our warehouse):
```{r tables, cache=FALSE}
membership <- get_powerschool("ps_membership_reg") %>% 
  select(studentid,
         schoolid,
         date = calendardate,
         enrolled = studentmembership,
         grade_level,
         attendance = ATT_CalcCntPresentAbsent,
         yearid) %>%
  filter(yearid >= 25)


attendance <- get_powerschool("attendance") %>% 
  filter(yearid >= 25,
        att_mode_code == "ATT_ModeDaily") 

attendance_code <- get_powerschool("attendance_code") %>%
  mutate(att_code = if_else(att_code == "true", "T", att_code)) %>% 
  select(attendance_codeid = id,
         att_code)
  
  
```
And now we do some pre-procecessing:
```{r joining tables, cache=TRUE}
member_att <- membership %>% 
  dplyr::left_join(attendance %>% 
              select(-schoolid,
                        -yearid) %>% 
              dplyr::left_join(attendance_code,
                        by = "attendance_codeid"),
            by = c("studentid",
                   "date" = "att_date")) %>% 
  collect()

```
Recoding enrollmeent and attendance data:
```{r recoding}
member_att %>% 
  janitor::tabyl(att_code)

student_att <- member_att %>%
  mutate(enrolled0 = 1,
         enrolled = if_else(att_code == "D" & !is.na(att_code), 0, enrolled0),
         present0 = ifelse(is.na(att_code), 1, 0),
         present1 = ifelse(att_code %in%  c("A", "S"), 0, present0),
         present2 = ifelse(att_code == "H", 0.5, present1),
         present3 = ifelse(att_code %in% c("T", "E", "I"), 1, present2),
         present = ifelse(is.na(present2), 1, present3),
         absent = (1 - present)*enrolled,
         tardy = ifelse(att_code %in% "T", 1, 0),
         dna0 = if_else(att_code == "D", 1, 0),
         dna = if_else(is.na(dna0), 0, dna0)) %>%
  select(yearid,
         schoolid,
         studentid,
         grade_level,
         date,
         att_code,
         enrolled,
         present,
         absent,
         tardy,
         dna)


```

Now we calculate some quantitities of interest: year_end ADA, cumulative ADA by date
```{r cumulative quantities of interest}
yearly_ada <- student_att %>% 
  filter(yearid < 28) %>%
  group_by(yearid,
           studentid) %>% 
  summarize(enrolled = sum(enrolled),
            absent = sum(absent),
            year_end = 1 - (absent/enrolled))
 
cum_ada <- student_att %>% 
  filter(yearid < 28) %>%
  group_by(yearid,
           studentid) %>% 
  arrange(date) %>% 
  mutate(cum_enrolled = cumsum(enrolled),
         cum_absent = cumsum(absent),
         running_ada = 1 - (cum_absent/cum_enrolled)) %>% 
  filter(cum_enrolled > 0)

yearly_cum_ada <- cum_ada %>% 
  select(yearid,
         schoolid,
         grade_level,
         studentid,
         date,
         cum_enrolled,
         cum_absent,
         running_ada) %>% 
  dplyr::left_join(yearly_ada %>% 
              select(-c(enrolled,
                        absent)),
            by = c("yearid",
                   "studentid"))

yearly_cum_ada %>% filter(studentid == 14852)
```

Now we begin doing some subsetting to build a basic model. We'll pull the data for each student over the last few years (excluding SY 18-19) for the last school day in December.  We'll build our model off of this "snapshot". The resulting table shows for each student their number of days enrolled (`cum_enrolled`), days absent (`cum_absent`), ADA through last school day in december (`running_ada`), and the given year's EOY ADA (`year_end`).
```{r dec_subset}
dec_dates <- yearly_cum_ada %>% 
  group_by(yearid) %>% 
  select(date) %>% 
  distinct() %>% 
  filter(lubridate::month(date) == 12) %>% 
  filter(date == max(date))

yca_filtered <- yearly_cum_ada %>% 
  inner_join(dec_dates %>% 
               ungroup() %>% 
               select(-yearid),
             by = c("date")) %>%
  ungroup() %>%
  mutate(yearid = as.factor(yearid))


yca_filtered
  
```
## Models 

our first model (`mod_0`) models year end ADA as a function of cumulative days enrolled and cumulaitve days absent through the last day of December, using school year as a fixed effect:

$$ y_{i} = f(\alpha + \beta_{e}d_{e} + \beta_{a}d_{a} +\delta_{year} ) $$

We'll first fit a linear model:
```{r lm}

formula_0 <- as.formula(year_end ~ cum_enrolled + cum_absent + yearid)

mod_0 <- lm(formula = formula_0, 
            data = yca_filtered)

summary(mod_0)

  
```
Let's plot the ranked residulas of `mod_0`)
```{r plot_residuals_lm}
# yca_filtered %>% 
#   ungroup() %>% 
#   modelr::add_predictions(mod_0) %>% 
#   modelr::add_residuals(mod_0) %>% 
#   arrange(resid) %>% 
#   group_by(yearid) %>% 
#   mutate(rank = row_number(resid)) %>% 
#   ggplot(aes(x = rank, y = resid)) +
#   geom_segment(aes(xend = rank, yend = 0),
#                size = .075, alph=.3) +
#   facet_grid(yearid~.)


plot_ranked_residuals <- function(data,model){
data %>% 
  modelr::add_predictions(model) %>% 
  modelr::add_residuals(model) %>% 
  arrange(resid) %>% 
  group_by(yearid) %>% 
  mutate(rank = row_number(resid)) %>% 
  ggplot(aes(x = rank, y = resid)) +
  geom_segment(aes(xend = rank, yend = 0),
               size = .075, alpha =  .3) +
  facet_grid(yearid~.)

    
}

plot_ranked_residuals(yca_filtered, mod_0)

```
And heres the residuals as a funtion of actuals. 
```{r}
# yca_filtered %>% 
#   ungroup() %>% 
#   modelr::add_predictions(mod_0) %>% 
#   modelr::add_residuals(mod_0) %>% 
#   arrange(resid) %>% 
#   group_by(yearid) %>% 
#   mutate(rank = row_number(resid)) %>% 
#   ggplot(aes(x = year_end, y = resid)) +
#   geom_point(aes(color = cum_enrolled),
#              size = .001) +
#   scale_color_kipp(discrete = FALSE) +
#   facet_grid(yearid ~.)

plot_actual_v_residuals <- function(data, model){
  data %>% 
  ungroup() %>% 
  modelr::add_predictions(model) %>% 
  modelr::add_residuals(model) %>% 
  arrange(resid) %>% 
  group_by(yearid) %>% 
  mutate(rank = row_number(resid)) %>% 
  ggplot(aes(x = year_end, y = resid)) +
  geom_point(aes(color = cum_enrolled),
             size = .001) +
  scale_color_kipp(discrete = FALSE) +
  facet_grid(yearid ~.)
}
  

plot_actual_v_residuals(yca_filtered, mod_0)
```
```{r predicted v actual}
plot_actual_v_predicted <- function(data, model){
  data %>% 
  ungroup() %>% 
  modelr::add_predictions(model) %>% 
  modelr::add_residuals(model) %>% 
  arrange(resid) %>% 
  group_by(yearid) %>% 
  mutate(rank = row_number(resid)) %>% 
  ggplot(aes(x = year_end, y = pred)) +
  geom_point(aes(color = cum_enrolled),
             size = .001) +
  scale_color_kipp(discrete = FALSE) +
  facet_grid(yearid ~.)
}

plot_actual_v_predicted(yca_filtered, mod_0)
```

Let's fits these with a logit
```{r logit}
mod_1 <- glm(formula = formula_0,
             family = binomial(link = "logit"),
             data = yca_filtered)

summary(mod_1)

plot_ranked_residuals(yca_filtered, mod_1)
plot_actual_v_residuals(yca_filtered, mod_1)

```
```{r}
yca_filtered %>% 
  ungroup() %>%
  modelr::add_predictions(mod_1) %>%
  modelr::add_residuals(mod_1)
```
## Partitioning 

WE need to partition in train and test to better evaluate our models.  

We'll take [rsample](tidymodels.github.io/rsample) for a spin!

```{r rsample}
yca_split <- yca_filtered %>%
  rsample::initial_split(prop = .75)

yca_train <- analysis(yca_split)
yca_test <- assessment(yca_split)


mod_lm_train <- lm(formula = formula_0, data = yca_train)

yca_augmented <- broom::augment(mod_lm_train, 
                                newdata = yca_test) %>%
  mutate(resid = year_end - .fitted)

yca_augmented %>% 
  ggplot(aes(x = year_end, y = .fitted)) +
  geom_point(aes(color = cum_enrolled, size = -cum_enrolled)) +
  geom_abline(aes(intercept = 0, slope = 1)) +
  facet_grid(yearid~.) +
  scale_color_kipp(discrete = FALSE)

```
```{r size by se}
yca_augmented %>% 
  ggplot(aes(x = year_end, y = .fitted)) +
  geom_point(aes(color = cum_enrolled, size = .se.fit, shape = cum_absent>=16)) +
  geom_abline(aes(intercept = 0, slope = 1)) +
  facet_grid(yearid~.) +
  scale_color_kipp(discrete = FALSE)
```

To do: we should distance to school.  
