---
title: "Equating as measurement error model"
author: "Chris Haid"
date: "April 22, 2014"
output: html_document
---
So let's try test equating as a measurement error model and let's get some comparisons with `Stan` and `JAGS`.  Let's assume that a student $i \in N$ has an underlying ability measured on NWEA's RIT scale $r_i>0 \forall i \in N$.  The MAP provides a noisy measure of $r$ that I'll denote $\hat{r}$, where $r = [r_{1}, r_{2}, \ldots, r_{N}]$ and similarly for $\hat{r}$.

We'll first sample the $r_i$ from some likely scores and then create a noisy measure of
each.  Finally we'll linearly relate the $r_i$ to anther noisy measure $\hat{a} \in [0,1]$.

So first let's sample $r$ from a known distribution
```{r sample_r}
set.seed(432104)
require(rstan)
N <- 30 # number of students

r<-rnorm(n = N, mean = 211.39, sd =5.83) #8th grade Math NWEA Norm
head(r)
```

Now $\hat{r}$ is noisily measured from $r$:

```{r rhats}
sigma.r<-1.2
rhat<-rnorm(n=N, mean=r, sd=sigma.r)

head(rhat)
```

And finally we simulate $\hat{a}$ from $r$ as well.  In this case I want $a$ to be on the on the $[0,1]$ scale and so I pick $\alpha$ and $\beta$ to normalize the $r$'s on to that scale linearly:


```{r ahats}
  alpha <- -min(r)/diff(range(r))
  beta <- 1/diff(range(r))

  alpha

  beta

  sigma.a=.01
  ahat <- rnorm(n = N, mean = alpha + beta*r, sd = sigma.a)
  ahat[ahat > 1] <- 1
  ahat[ahat < 0] <- 0

head(ahat)
```

So we our data are the observed $\hat{r}_i$'s and $\hat{a}_i$'s. Our parameters of interest are the $r$, $\alpha$, and $\beta$.  I take the standard deviations---$\sigma_{r}$ and $\sigma_a$ to be unknown hyper parameters. So here's a `Stan` model that should estimate these values:

```{r stan_measurement_model}
model.1<-'
  data { 
    int<lower=0> N;
    real<lower=0,upper=1> ahat[N];
    real<lower=190, upper=240> rhat[N];
  }
  parameters{
    real alpha;
    real beta;
    real<lower=190, upper=240> r[N];
    real<lower=0> sigma_r;
    real<lower=0> sigma_a;
  }
  model{
    rhat ~ normal(r, sigma_r);
    for (i in 1:N) {
      ahat[i] ~ normal(alpha + beta * r[i], sigma_a) T[0,1];
    }
    alpha ~ normal(0,10);
    beta ~ normal(0,10);
    sigma_r ~ cauchy(0,5);
    sigma_a ~ cauchy(0,5);
  }
'
```

And now for the sampling:

```{r model1}
stan.data <- list('ahat' = ahat,
                  'rhat' = rhat,
                  'N' = N)

fit1 <- stan(model_code = model.1,
                   data = stan.data,
                   #pars = c('a','r', 'alpha', 'beta'),
                   chains=2,
                   iter=100)


fit2 <- stan(fit = fit1,
             data = stan.data,
             #pars = c(,'r', 'alpha', 'beta'),
             chains=4,
             iter=10000)

fit2

plot(fit2)
rstan::traceplot(fit2)
```

