---
title: "census_and_students"
params: 
  inputDir: "gathered/"
  thisDir:  "cook/"
  outputDir: "cooked/"
  fileStem: "census_and_students"
  inputFiles: 
    - "Census_Demographics"
    - "students"
---

```{r KnitrOpts, echo=FALSE}
knitr::opts_knit$set(root.dir = normalizePath('../'))
knitr::opts_chunk$set(echo=TRUE, message=TRUE, warning=TRUE, error=TRUE)
```

```{r setup}
library(tidyverse)
library(stringr)
library(sf)
library(kippcolors)
```


```{r Load, results='hide'}
loadedObjects <- represtools::LoadObjects(params)
```

```{r clean_students}
glimpse(students)

students_geocoded <- students %>% 
  filter(!is.na(geocode)) %>%
  separate(geocode, sep = ", ", into= c("lat", "lon")) %>%
  mutate(lat = as.numeric(str_replace(lat, "Lat: ", "")),
         lon = as.numeric(str_replace(lon, "Lng: ", ""))
         )
  
  
```

This is quick plot to show that geocoded students are being plotted.
```{r test_plot}
ggplot(students_geocoded) +
  geom_sf(data=chicago_tracts) +
  geom_point(aes(x = lon, y=lat), color = "hotpink") +
  facet_wrap(~ethnicity)


```
Google's geocoding isn't totally accurate.  For example these students have the wrong lat/lon given their actual addressess
```{r}
students_geocoded %>% filter(lon <= -87.8)
```

Create simple features data.frame form `students` and reduce to only those students with lat/lon within Chicago's borders
```{r students_sf}
students_sf <- students_geocoded %>%
  sf::st_as_sf(coords = c("lon", "lat"))

st_crs(students_sf)

st_crs(students_sf) <- st_crs(chi)

students_within<-st_within(students_sf, chi)

students_mask <- students_within %>% 
  map_lgl(function(x) {
  if(length(x) == 1) {
    TRUE
  } else {
    FALSE
  }
})


students_sf<-students_sf[students_mask,]

# ggplot(students_sf) +
#   geom_sf(data=chicago) +
#   geom_sf( color = "hotpink") +
#   facet_wrap(~ethnicity)
  
```

Recode student ethnicities from integer to string and schools from schoolid to actual name.
```{r ethnicities}
students_sf <- students_sf %>%
  mutate(ethnicity_name = case_when(
    ethnicity == 1 ~ "White",
    ethnicity == 2 ~ "African American",
    ethnicity == 4 ~ "Asian",
    ethnicity == 5 ~ "Hispanic",
    ethnicity == 6 ~ "Native Hawaiian/Pacific Islender",
    ethnicity == 9 ~ "Mulit-racial",
    ethnicity == 11 ~ "American Indian/Alaskan Native"),
    
    school = case_when(
      schoolid == 400180  ~ "KOA",
      schoolid == 4001802 ~ "KOP")
      
    )
```

Add ethnicity name to to Chicago data
```{r chicago_percents}
chicago_tracts <- chicago_tracts %>%
  mutate(percent = estimate/summary_est*100,
         ethnicity_name = case_when(
           variable == "total_white" ~ "White",
           variable == "total_aa" ~ "African American",
           variable == "total_hisp" ~ "Hispanic"
         ))

# mapview::mapview(chicago_tracts %>%
#                    filter(ethnicity_name == "Hispanic"),
#                  zcol = "percent") %>%
#   mapview::addFeatures(data = students_sf, zcol = "ethnicity_names")
```

Geocode KIPP One Location.  Get google terrain map for background.  Get student distances.
```{r kipp_one_location}

kipp_one_location <- ggmap::geocode("KIPP One Academy, Chicago")

one_sf <- kipp_one_location %>%
  mutate(school_label = "KIPP One") %>%
  sf::st_as_sf(coords = c("lon", "lat"), remove = FALSE)

chi_bg_map <- ggmap::get_map(location = kipp_one_location,
                                   zoom = 13,
                                   maptype = "terrain",
                                   source = "google",
                                   color = "bw"
                                   )


# Set CRS
st_crs(one_sf) <- st_crs(students_sf)

chi_map <- ggmap::ggmap(chi_bg_map)

chi_map +
  coord_sf() +
  geom_sf(data = chicago_tracts %>%
            filter(ethnicity_name %in% c("African American", "Hispanic")),
          aes(fill = percent), color = NA, alpha = .3,
          inherit.aes = FALSE) +
  geom_sf(data = students_sf %>%
            filter(ethnicity_name %in% c("African American", "Hispanic")), 
          aes(color = ethnicity_name),
          size = 2,
          alpha = .8,
          inherit.aes = FALSE) +
    geom_sf(data = one_sf,
          size = 2,
          shape = 13,
          color = "darkblue",
          inherit.aes = FALSE) +
  ggrepel::geom_text_repel(data = one_sf,
             aes(x = lon, y=lat, label = school_label),
             color = "darkblue",
             size = 2) +
  scale_fill_viridis_c() +
  #scale_fill_viridis_d()
  #facet_grid(grade_level~ethnicity_name)
  facet_grid(school~ethnicity_name)
```

Sample points in census tracts
```{r st_sample}

distinct_race<-chicago_tracts %>%
  as_data_frame() %>%
  select(variable) %>%
  distinct()

#function to sample, recast as sf

sample_points_as_sf <- function(ethnicity, .data = chicago_tracts, pop_per_point = 10) {
  
  filtered_df <- .data %>%
    filter(variable == ethnicity)
  
  geometry <- filtered_df %>%
    st_sample(size = filtered_df$estimate/pop_per_point)
  
  #coerce to sf
  out_sf <- st_sf(geometry)
  
  # Add Ethnicity column
  out_sf <- out_sf %>%
    mutate(variable = ethnicity) %>%
    mutate(ethnicity_name = case_when(
           variable == "total_white" ~ "White",
           variable == "total_aa" ~ "African American",
           variable == "total_hisp" ~ "Hispanic")) #%>%
  #rename(geometry = sampled_sfc)
  
  # Return 
  out_sf

}


# x <- sample_points_as_sf("total_aa", pop_per_point = 10000)
# glimpse(x)

sampled_points_sf <- distinct_race$variable %>%
  map(~ sample_points_as_sf(ethnicity = ., 
                               .data = chicago_tracts,
                               pop_per_point = 300)) %>%
  reduce(rbind)

  #scale_fill_viridis_d()
  #facet_grid(grade_level~ethnicity_name)
  #facet_grid(school~ethnicity_name)
```
Distances form KOA
```{r}
km_to_miles <- function(x) x * 0.621371

sampled_points_dist <- km_to_miles(as.integer(st_distance(one_sf, sampled_points_sf)/1000))
sampled_points_sf$distance <- sampled_points_dist

students_dist <- km_to_miles(as.integer(st_distance(one_sf, students_sf)/1000))
students_sf$distance <- students_dist
# sample_points_sf <- sampled_points_list %>%
#   map_df(~bind_cols(.))
  
#   
# aa<-chicago_tracts %>%
#   filter(variable == "total_aa") 
# 
# x_aa <- aa %>% st_sample(size = aa$estimate/300)
# 
# 
# hisp <- aa<-chicago_tracts %>%
#   filter(variable == "total_hisp") 
# 
# x_hisp <- hisp %>% st_sample(size = hisp$estimate/300)
# 
# sf::as
# 
# chicago_tracts_sampled <- chicago_tracts %>%
#  
# z <- chicago_tracts %>%
#   mutate(sampled = st_sample(chicago_tracts$geometry, size = chicago_tracts$estimate/1000))

chi_map +
  geom_sf(data = sampled_points_sf %>% filter(distance <= 1.5), 
          aes(color = ethnicity_name), 
          alpha = .7,
          inherit.aes = FALSE) +
    geom_sf(data = one_sf,
          size = 2,
          shape = 13,
          color = "darkblue",
          inherit.aes = FALSE) +
  ggrepel::geom_text_repel(data = one_sf,
             aes(x = lon, y=lat, label = school_label),
             color = "darkblue",
             size = 2) +
  scale_color_kipp(reverse= TRUE) #+
```

```{r ListObjects}
lstObjects <- represtools::ListObjects(pattern = c("_sf", "_map", "_tracts"))
represtools::DescribeObjects(represtools::NamesToObjects(lstObjects))
```

```{r Save}
save(file = represtools::OutputFile(params)
     , list = lstObjects)
```
