MAP Analysis of KAPS 
========================================================

Bubble Charts for Ellen
------------------------------------------------------------------------------

```{r load_packages}
library(reshape)
library(ggplot2)
library(grid)
library(RJDBC)
library(plyr)
library(data.table)

setwd("/Users/chaid/Dropbox/Consulting/KIPP Ascend/Testing Analysis/MAP/Analysis/KAP")

```
```{r db_connections, echo=FALSE, cache=FALSE}

setwd('/Users/chaid/Dropbox/Consulting/KIPP Ascend/Testing Analysis/MAP/Analysis')

drvr <- JDBC("com.mysql.jdbc.Driver","/Users/chaid/Dropbox/JDBC Drivers/mysql-connector-java-5.0.8/mysql-connector-java-5.0.8-bin.jar","")

con <- dbConnect(drvr,"jdbc:mysql://54.245.118.235/db_kippchidata", "chaid", "haiKIPP1" )

```

```{r Primary_Bubble_Chart_data, echo=FALSE, cache=TRUE}



map.1213<-dbGetQuery(con, 'CALL GetMAPResultsFromTo("Fall12","Winter13");')
map.1213<-data.table(map.1213)

#####F11-W12
map.1112<-dbGetQuery(con, 'CALL GetMAPResultsFromTo("Fall11","Spring12");')
map.1112<-data.table(map.1112)

#Add Years to two tables
map.1112[,Year:=2012]
map.1213[,Year:=2013]

#Change names of columns for test data in prep for cbind

setnames(map.1112, names(map.1112), sub("Fall11", "Fall", names(map.1112)))
setnames(map.1213, names(map.1213), sub("Fall12", "Fall", names(map.1213)))
setnames(map.1112, names(map.1112), sub("Winter12", "Winter", names(map.1112)))
setnames(map.1213, names(map.1213), sub("Winter13", "Winter", names(map.1213)))

map.scores<-rbindlist(list(map.1112,map.1213))

#map.scores[Grade==13,Grade:=0]

#Check contents
map.scores

#Reorder levels (since 13=Kinder, prior to Fall 2012, after that it is Kinder=0) and rename
map.scores$Fall_Grade <- factor(map.scores$Fall_Grade, levels=c("0", "1","2", "5", "6","7","8"))
levels(map.scores$Fall_Grade) <- c("K", "1", "2", "5", "6","7","8")

map.scores.by.grade<-ddply(map.scores[,c("ID","Fall_Grade","ClassName","Subject","Fall_RIT", "Winter_RIT","TypicalFallToWinterGrowth", "Year"), with=FALSE],
                           .(Subject, Fall_Grade, Year), function(df) orderid(df,df$Fall_RIT))
map.scores.by.class<-ddply(map.scores[,c("StudentID","Fall_Grade","ClassName","MeasurementScale","RIT_Fall", "RIT_Winter", "TypicalFallToWinterGrowth"), with=FALSE],
                           .(MeasurementScale, ClassName), function(df) orderid(df,df$RIT_Fall))



primary<-c("K","1", "2")
map.k1<-map.scores[Fall_Grade %in% primary]
map.k1[Fall_Pctl<25, Quartile:=1]
map.k1[Fall_Pctl>=25&map.k1$Fall_Pctl<50, Quartile:=2]
map.k1[Fall_Pctl>=50&map.k1$Fall_Pctl<75, Quartile:=3]
map.k1[Fall_Pctl>=75, Quartile:=4]

map.k1$MetExpGrowth<-map.k1$Fall_RIT+map.k1$TypicalFallToWinterGrowth<=map.k1$Winter_RIT

k1.prop<-ddply(map.k1, .(Quartile, Fall_Grade, Subject, Year), summarize, pr= sum(MetExpGrowth)/length(MetExpGrowth), NumTested=length(MetExpGrowth), NumMetGrowth=sum(MetExpGrowth))



KCCP.grades<-c("5")
map.kccp<-map.scores[Fall_Grade %in% KCCP.grades & SchoolName=="KIPP Create Middle School"]
map.kccp$Quartile[map.kccp$Fall12_Pctl<25]<-1
map.kccp$Quartile[map.kccp$Fall12_Pctl>=25&map.kccp$Fall12_Pctl<50]<-2
map.kccp$Quartile[map.kccp$Fall12_Pctl>=50&map.kccp$Fall12_Pctl<75]<-3
map.kccp$Quartile[map.kccp$Fall12_Pctl>=75]<-4
map.kccp$MetExpGrowth<-map.kccp$Fall12_RIT+map.kccp$TypicalFallToWinterGrowth<=map.kccp$Winter13_RIT

kccp.prop<-ddply(map.kccp, .(Quartile, Fall_Grade, Subject, Year), summarize, pr= sum(MetExpGrowth)/length(MetExpGrowth), NumTested=length(MetExpGrowth), NumMetGrowth=sum(MetExpGrowth))


KAMS.grades<-c("7", "8")
map.kams<-map.scores[Fall_Grade %in% KAMS.grades & SchoolName=="KIPP Ascend Middle School"]

#map.kams<-as.data.frame(map.kams)
map.kams$Quartile[map.kams$Fall_Pctl<25]<-1
map.kams$Quartile[map.kams$Fall_Pctl>=25&map.kams$Fall_Pctl<50]<-2
map.kams$Quartile[map.kams$Fall_Pctl>=50&map.kams$Fall_Pctl<75]<-3
map.kams$Quartile[map.kams$Fall_Pctl>=75]<-4
map.kams$MetExpGrowth<-map.kams$Fall_RIT+map.kams$TypicalFallToWinterGrowth<=map.kams$Winter_RIT

kams.prop<-ddply(map.kams, .(Quartile, Fall_Grade, Subject, Year), summarize, pr= sum(MetExpGrowth)/length(MetExpGrowth), NumTested=length(MetExpGrowth), NumMetGrowth=sum(MetExpGrowth))

```

```{r KAPS_KAMS_KCCP_Bubble_Chart, echo=FALSE, fig.height=10, fig.width=10}
p.hunch <- ggplot(k1.prop,aes(x=Quartile, y=pr, color =Fall_Grade), stat="identity") +
  geom_point(aes(size=NumTested, alpha=as.factor(Year))) + 
  scale_area(range = c(8,30), guide="none") + 
  geom_text(aes(x=Quartile, y=pr, label=paste(round(pr,2)*100,"%",sep=""), group=Fall_Grade),
            color='black',size=3.5,vjust=-.5) + 
  geom_text(aes(x=Quartile, y=pr, label=paste("(",NumMetGrowth,"/", NumTested,")",sep=''
                                              )
                ), color='black', size=2, vjust=1) + 
  scale_x_continuous("", breaks=c(1,2,3,4), labels=c("1st\nQuartile","2nd\nQuartile",
                                                     "3rd\nQuartile", "4th\nQuartile")) +
  scale_y_continuous("% Exceeding Typical Growth") +
  scale_alpha_discrete("",range=c(.5,.9), guide="none") +
  scale_color_discrete(guide="none") +
  expand_limits(x=c(0,5), y=c(-.1,1.1)) + 
  facet_grid(Grade~Subject) + 
  ggtitle("KIPP Primary MAP  Meets/Exceeds Growth by Fall Quartile: F11-W12 vs F12-W13") + 
  theme_bw() +
  theme(legend.key.size = unit(1, "cm"), axis.text.x=element_text(size=8))
p.hunch

ggsave(p.hunch,file="MAP_KAPS_W13_Bubble_130222.pdf", width=11, height=8.5)


```{r Primary_Bubble_Chart, echo=FALSE, fig.height=10, fig.width=10}
p.kccp <- ggplot(kccp.prop,aes(x=Quartile, y=pr, color =Fall_Grade), stat="identity") +
  geom_point(aes(size=NumTested)) + 
  scale_area(range = c(8,30)) + 
  geom_text(aes(x=Quartile, y=pr, label=paste(round(pr,2)*100,"%",sep=""), group=Fall_Grade),
            color='black',size=3.5,vjust=-.5) + 
  geom_text(aes(x=Quartile, y=pr, label=paste("(",NumMetGrowth,"/", NumTested,")",sep=''
                                              )
                ), color='black', size=2, vjust=1) + 
  scale_x_continuous("", breaks=c(1,2,3,4), labels=c("1st\nQuartile","2nd\nQuartile",
                                                     "3rd\nQuartile", "4th\nQuartile")) +
  scale_y_continuous("% Exceeding Typical Growth") +
  expand_limits(x=c(0,5), y=c(-.1,1.1)) + 
  facet_grid(Fall_Grade~Subject) + 
  ggtitle("KCCP MAP F12-W13 Meets/Exceeds Growth by Fall 2012 Quartile") + 
  theme_bw() +
  theme(legend.key.size = unit(1, "cm"))
p.kccp

ggsave(p.kccp,file="MAP_KCCP_W13_Bubble.pdf", width=11, height=8.5)


p.kams <- ggplot(kams.prop,aes(x=Quartile, y=pr, color =Fall_Grade), stat="identity") +
  geom_point(aes(size=NumTested)) + 
  scale_area(range = c(8,30)) + 
  geom_text(aes(x=Quartile, y=pr, label=paste(round(pr,2)*100,"%",sep=""), group=Fall_Grade),
            color='black',size=3.5,vjust=-.5) + 
  geom_text(aes(x=Quartile, y=pr, label=paste("(",NumMetGrowth,"/", NumTested,")",sep=''
                                              )
                ), color='black', size=2, vjust=1) + 
  scale_x_continuous("", breaks=c(1,2,3,4), labels=c("1st\nQuartile","2nd\nQuartile",
                                                     "3rd\nQuartile", "4th\nQuartile")) +
  scale_y_continuous("% Exceeding Typical Growth") +
  expand_limits(x=c(0,5), y=c(-.1,1.1)) + 
  facet_grid(Fall_Grade~Subject) + 
  ggtitle("KAMS MAP F12-W13 Meets/Exceeds Growth by Fall 2012 Quartile") + 
  theme_bw() +
  theme(legend.key.size = unit(1, "cm"))
p.kams


p.kams <- ggplot(kams.prop,aes(x=Quartile, y=pr, color =Fall_Grade), stat="identity") +
  geom_point(aes(size=NumTested, alpha=as.factor(Year))) + 
  scale_area(range = c(8,30), guide="none") + 
  geom_text(aes(x=Quartile, y=pr, label=paste(round(pr,2)*100,"%",sep=""), group=Fall_Grade),
            color='black',size=3.5,vjust=-.5) + 
  geom_text(aes(x=Quartile, y=pr, label=paste("(",NumMetGrowth,"/", NumTested,")",sep=''
                                              )
                ), color='black', size=2, vjust=1) + 
  scale_x_continuous("", breaks=c(1,2,3,4), labels=c("1st\nQuartile","2nd\nQuartile",
                                                     "3rd\nQuartile", "4th\nQuartile")) +
  scale_y_continuous("% Exceeding Typical Growth") +
  scale_alpha_discrete("",range=c(.5,.9), guide="none") +
  scale_color_discrete(guide="none") +
  expand_limits(x=c(0,5), y=c(-.1,1.1)) + 
  facet_grid(Grade~Subject) + 
  ggtitle("KIPP Primary MAP  Meets/Exceeds Growth by Fall Quartile: F11-W12 vs F12-W13") + 
  theme_bw() +
  theme(legend.key.size = unit(1, "cm"), axis.text.x=element_text(size=8))
p.hunch


ggsave(p.kams,file="MAP_KAMS_W13_Bubble.pdf", width=11, height=8.5)


```