---
title: "TNTP Fall 2018 Silo-prep script"
output: html_notebook
params:
  region: "KIPP Chicago"
---


```{r packages}
library(tidyverse)
library(tntpr)
library(janitor)

```

Let's read in the four (4!) new TNTP files that we download form their [Tableau site](https://data.tntp.org).

Note that you need to use read.delim b/c of encoding issues
```{r extract}

tntp_item_benchmarks <- read.delim("../data/2018-2019/SS_Benchmarks_data_fall_16_17_18.csv", sep="\t",
                              fileEncoding = "UTF-16LE", 
                              stringsAsFactors = FALSE
                              ) %>%
  clean_names %>%
  as_tibble()


tntp_domains_benchmarks <- read.delim("../data/2018-2019/SS_benchmarks_for_Domains_data_fall_16_17_18.csv", 
                                      sep="\t",
                              fileEncoding = "UTF-16LE", 
                              stringsAsFactors = FALSE
                              ) %>%
  clean_names %>%
  as_tibble()


tntp_domains <- read.delim("../data/2018-2019/SS_Domains_data_fall_16_17_18.csv", 
                                      sep="\t",
                              fileEncoding = "UTF-16LE", 
                              stringsAsFactors = FALSE
                              ) %>%
  clean_names %>%
  as_tibble()


tntp_items <- read.delim("../data/2018-2019/SS_Items_data_fall_16_17_18.csv", 
                                      sep="\t",
                              fileEncoding = "UTF-16LE", 
                              stringsAsFactors = FALSE
                              ) %>%
  clean_names %>%
  as_tibble()



```

Need to align column naming conventions, rename "All Schools Below" to Region name (`params$region`), remove astericks from questions.

## Domains
Break down by, meaure_names -> variable
transformed_score, measure_values -> value

```{r transform_domains}

domains_out <- tntp_domains %>%
  rename(variable = break_down_by, 
         value = transformed_score
           )

domains_benchmarks_out <- tntp_domains_benchmarks %>%
  rename(variable = measure_names, 
         value = measure_values
         ) %>%
  dplyr::mutate(variable = if_else(variable == "Average of schools below", params$region, variable))

```

## Items
Break down by, meaure_names -> variable
transformed_score, measure_values -> value
```{r transform_items}
items_out <- tntp_items %>%
  rename(variable = break_down_by, 
         value = transformed_score
           ) %>%
  mutate(clean_question = str_remove_all(clean_question, "\\*?")) %>%
  select(-school_header)

items_benchmarks_out <- tntp_item_benchmarks %>%
  rename(variable = measure_names, 
         value = measure_values
         ) %>%
  dplyr::mutate(variable = if_else(variable == "Average of schools below", params$region, variable),
                clean_question = str_remove_all(clean_question, "\\*?"))
```

## Load (sorta)
```{r load}
write_csv(items_out,"../out_for_SILO/items_fall_16_17_18.csv")
write_csv(items_benchmarks_out,"../out_for_SILO/items_benchmarks_fall_16_17_18.csv")
write_csv(domains_out,"../out_for_SILO/domains_fall_16_17_18.csv")
write_csv(domains_benchmarks_out,"../out_for_SILO/domains_benchmarks_fall_16_17_18.csv")

```

