 KIPP Chicago Regional Attendance Report
========================================================
```{r options, echo=FALSE}

#knit2html("Weekly_Attendance_Report_131003.Rmd", output="test.html", options=c('fragment_only', 'skip_style')); system('cp ./test.html ~/Sites/test.local/'); system('cp -r ./figure/* ~/Sites/test.local/figure/')


#source('.Rprofile')
options(#replace.assign=TRUE,
        width=60
        )
opts_knit$set(
   self.contained=FALSE
  )
opts_chunk$set( 
               #fig.align='center', 
               #dev='tikz', 
               #fig.width=8, 
               #fig.height=10.5, 
               #fig.show='hold', 
               cache=FALSE, 
               par=TRUE,
               echo=FALSE,
               message=FALSE,
               warning=FALSE,
               autodepend=FALSE
               )
               
```


```{r load_libraries}


library(plyr)  #To manipulate data
library(reshape) #More data manipulation
library(ggplot2) #Graphics of grammer graphing
library(grid) #More Graphing
library(gridExtra) #better than par(mfrow=c(r,c)) for arranging ggplot2 and lattice graphics
library(lubridate) #for dealing with dates
library(xtable) #for pretty tables
library(RJDBC) #To get data form PowerSchool DB
library(data.table)
#library(RODBC) #To get data form PowerSchool DB
#source MAP helper functions
source("~/Dropbox/Consulting/KIPP Ascend/Data Analysis/MAP/Code/R/MAP_helper_functions.R")
```
```{r establsih_DB_connections, cache=FALSE}
drvr <- JDBC("oracle.jdbc.driver.OracleDriver", "/Users/chaid/Dropbox/JDBC Drivers/ojdbc6.jar","")
drvr <- JDBC("oracle.jdbc.driver.OracleDriver", "/Users/chaid/Dropbox/JDBC Drivers/ojdbc6.jar","")

pscon <- dbConnect(drvr,"jdbc:oracle:thin:@10.160.29.47:1521:IL039","psnavigator","laidephy")
```

```{r calculate_date, cach=FALSE}
date.first  <- "2013-08-19" # first day of school year
date.second <- floor_date(today(), unit="week") # 
date.title <- date.second -weeks(1) + days(1) 
title.date<-paste(lubridate::month(date.title, 
                                   label=TRUE, 
                                   abbr=FALSE),
                  " ",
                  lubridate::day(date.title), ", ", 
                  lubridate::year(date.title), sep=""
                  )
```
Week of `r I(title.date)`
--------------------------------------------------------
### Attendance

```{r get_Attendance_Data, cache=TRUE}
#Get Attendance data.  
# This SQL statement retrieves each enrolled student and their Attendance status form PowerSchool 
# by date for a given dateThi sis a table with each student for each day with status for the day





sql.statement<-paste("SELECT 
    m.schoolid, 
    m.grade_level,
  	m.calendardate, 
	m.STUDENT_NUMBER AS StudentID, 
  	m.lastfirst,
	m.Enrolled,
	a.Att_Code,
	a.Description as AttDescr,
	a.Presence_Status_CD,
	a.COURSE_CREDIT_POINTS,
	CASE 
		WHEN	a.Presence_Status_CD = 'Absent' THEN a.COURSE_CREDIT_POINTS 
		ELSE 0 
	END as Absent
FROM (
	SELECT
		psmd.SchoolID,  
		psmd.grade_level, 
		psmd.calendardate, 
		psmd.studentid,
		s.STUDENT_NUMBER,
		s.LASTFIRST, 
		1 as Enrolled  
	FROM PS_Membership_Defaults psmd
	LEFT JOIN students s ON psmd.StudentID = s.id
	Where 	calendardate >= TO_DATE('",date.first,"','yyyy-mm-dd')
		AND  calendardate <= TO_DATE('",date.second,"','yyyy-mm-dd')
) m
LEFT JOIN (
	SELECT 
		att.schoolid, 
		att.StudentID,
		att.Att_Date,
		attc.Att_Code,
		attc.Description,
		attc.Presence_Status_CD,
		attc.COURSE_CREDIT_POINTS
	FROM Attendance att
  	INNER JOIN Attendance_Code attc ON att.Attendance_CodeID = attc.ID
	WHERE 
		att.Att_Mode_Code = 'ATT_ModeDaily'
  		AND att.Att_Date >= TO_DATE('",date.first,"','yyyy-mm-dd')
  		AND att.Att_Date <= TO_DATE('",date.second,"','yyyy-mm-dd')
		AND (attc.att_code = 'A' OR attc.att_code = 'S' or attc.att_code = 'X' or attc.att_code = 'H')
) a
ON m.STUDENTID = a.studentid AND m.calendardate =a.Att_Date AND m.schoolID = a.schoolid
ORDER BY schoolid, grade_level, calendardate",sep="")
  
  
  
  
Attendance<-dbGetQuery(pscon, sql.statement)
```
```{r polish_Attendance, cache=FALSE}

Attendance$SchoolInitials[Attendance$SCHOOLID==7810]<-"KAMS"
Attendance$SchoolInitials[Attendance$SCHOOLID==78102]<-"KAPS"
Attendance$SchoolInitials[Attendance$SCHOOLID==400146]<-"KCCP"
Attendance$SchoolInitials[Attendance$SCHOOLID==400163]<-"KBCP"
Attendance$SchoolInitials<-factor(Attendance$SchoolInitials, levels=c("KAPS", "KAMS", "KCCP", "KBCP"))


#Summarize enrollments and absences by day

DailyEnrollAttend<-ddply(Attendance, .(SCHOOLID, CALENDARDATE), summarise, Enrolled=sum(ENROLLED), Absent=sum(ABSENT))

DailyEnrollAttendByGrade<-ddply(Attendance, .(SCHOOLID, GRADE_LEVEL, CALENDARDATE), summarise, Enrolled=sum(ENROLLED), Absent=sum(ABSENT))
#rename some columns (becuase ALL CAPS is annoying)

names(DailyEnrollAttend)<-c("SchoolID", "Date", "Enrolled", "Absent")
names(DailyEnrollAttendByGrade)<-c("SchoolID", "Grade","Date", "Enrolled", "Absent")


#transform dates
DailyEnrollAttend$Date<-ymd_hms(DailyEnrollAttend$Date)
DailyEnrollAttendByGrade$Date<-ymd_hms(DailyEnrollAttendByGrade$Date)


#Some quick daily stats
DailyEnrollAttend$Present<-DailyEnrollAttend$Enrolled-DailyEnrollAttend$Absent
DailyEnrollAttendByGrade$Present<-DailyEnrollAttendByGrade$Enrolled-DailyEnrollAttendByGrade$Absent

DailyEnrollAttend$PctAbsent<-DailyEnrollAttend$Absent/DailyEnrollAttend$Enrolled
DailyEnrollAttendByGrade$PctAbsent<-DailyEnrollAttendByGrade$Absent/DailyEnrollAttendByGrade$Enrolled

DailyEnrollAttend$PctPresent<-1-DailyEnrollAttend$PctAbsent
DailyEnrollAttendByGrade$PctPresent<-1-DailyEnrollAttendByGrade$PctAbsent

DailyEnrollAttend$PctPresentGTE95<-DailyEnrollAttend$PctPresent>=.95
DailyEnrollAttendByGrade$PctPresentGTE95<-DailyEnrollAttendByGrade$PctPresent>=.95

DailyEnrollAttend$Present96PctBenchMark<-.96*DailyEnrollAttend$Enrolled
DailyEnrollAttendByGrade$Present96PctBenchMark<-.96*DailyEnrollAttendByGrade$Enrolled

DailyEnrollAttend$WeekInYear<-week(DailyEnrollAttend$Date)
DailyEnrollAttendByGrade$WeekInYear<-week(DailyEnrollAttendByGrade$Date)

DailyEnrollAttend$WeekInSchoolYear<-(intv<-floor_date(DailyEnrollAttend$Date, unit="week")-min(floor_date(DailyEnrollAttend$Date, unit="week")))/dweeks(1)+1
DailyEnrollAttendByGrade$WeekInSchoolYear<-DailyEnrollAttendByGrade$WeekInYear - min(DailyEnrollAttendByGrade$WeekInYear) +1

DailyEnrollAttend$WeekOfDate<-floor_date(DailyEnrollAttend$Date, unit="week") + days(1) 
DailyEnrollAttendByGrade$WeekOfDate<-floor_date(DailyEnrollAttendByGrade$Date, unit="week") + days(1) 

#Long Week (of) label
DailyEnrollAttend$WeekOfDateLabel<-paste("Week of \n", lubridate::month(DailyEnrollAttend$WeekOfDate,label=TRUE, abbr=TRUE), day(DailyEnrollAttend$WeekOfDate), sep=" ")
DailyEnrollAttendByGrade$WeekOfDateLabel<-paste("Week of \n", lubridate::month(DailyEnrollAttendByGrade$WeekOfDate,label=TRUE, abbr=TRUE), day(DailyEnrollAttendByGrade$WeekOfDate), sep=" ")

DailyEnrollAttend$WeekOfDateLabel<-factor(DailyEnrollAttend$WeekInSchoolYear, labels=unique(DailyEnrollAttend$WeekOfDateLabel))
DailyEnrollAttendByGrade$WeekOfDateLabel<-factor(DailyEnrollAttendByGrade$WeekInSchoolYear, labels=unique(DailyEnrollAttendByGrade$WeekOfDateLabel))

#Short Week  Label
DailyEnrollAttend$WeekOfShortDateLabel<-paste(lubridate::month(DailyEnrollAttend$WeekOfDate,label=TRUE, abbr=TRUE), day(DailyEnrollAttend$WeekOfDate), sep=" ")
DailyEnrollAttendByGrade$WeekOfShortDateLabel<-paste(lubridate::month(DailyEnrollAttendByGrade$WeekOfDate,label=TRUE, abbr=TRUE), day(DailyEnrollAttendByGrade$WeekOfDate), sep=" ")


DailyEnrollAttend$WeekOfShortDateLabel<-factor(DailyEnrollAttend$WeekInSchoolYear, labels=unique(DailyEnrollAttend$WeekOfShortDateLabel))
DailyEnrollAttendByGrade$WeekOfShortDateLabel<-factor(DailyEnrollAttendByGrade$WeekInSchoolYear, labels=unique(DailyEnrollAttendByGrade$WeekOfShortDateLabel))


#add School Initials for graphics
DailyEnrollAttend$SchoolInitials[DailyEnrollAttend$SchoolID==7810]<-"KAMS"
DailyEnrollAttendByGrade$SchoolInitials[DailyEnrollAttendByGrade$SchoolID==7810]<-"KAMS"

DailyEnrollAttend$SchoolInitials[DailyEnrollAttend$SchoolID==78102]<-"KAPS"
DailyEnrollAttendByGrade$SchoolInitials[DailyEnrollAttendByGrade$SchoolID==78102]<-"KAPS"

DailyEnrollAttend$SchoolInitials[DailyEnrollAttend$SchoolID==400146]<-"KCCP"
DailyEnrollAttendByGrade$SchoolInitials[DailyEnrollAttendByGrade$SchoolID==400146]<-"KCCP"

DailyEnrollAttend$SchoolInitials[DailyEnrollAttend$SchoolID==400163]<-"KBCP"
DailyEnrollAttendByGrade$SchoolInitials[DailyEnrollAttendByGrade$SchoolID==400163]<-"KBCP"

DailyEnrollAttend$SchoolInitials<-factor(DailyEnrollAttend$SchoolInitials, levels=c("KAPS", "KAMS", "KCCP", "KBCP"))
DailyEnrollAttendByGrade$SchoolInitials<-factor(DailyEnrollAttendByGrade$SchoolInitials, levels=c("KAPS", "KAMS", "KCCP", "KBCP"))
#head(DailyEnrollAttend)
```
```{r Attendance_Summaries}
#calculate Attendance rate by week by school
AttRateByWeekBySchool<-ddply(DailyEnrollAttend,.(SchoolInitials,WeekOfShortDateLabel), summarise, AttRate=sum(Present)/sum(Enrolled)*100)

AttRateYTDBySchool<-ddply(DailyEnrollAttend,.(SchoolInitials), summarise, AttRate=sum(Present)/sum(Enrolled)*100)

AttRateYTDBySchool$WeekOfShortLabel<-"YTD Each School"

AttRateByWeekBySchool.table<-cast(AttRateByWeekBySchool, WeekOfShortDateLabel ~ SchoolInitials)

AttRateYTDBySchool<-reshape(AttRateYTDBySchool, idvar="WeekOfShortLabel",timevar="SchoolInitials", direction="wide")
names(AttRateYTDBySchool)<-c("WeekOfShortDateLabel", "KAPS", "KAMS", "KCCP", "KBCP")




AttRateYTDRegion<-data.frame(WeekOfShortDateLabel="YTD Region", KAPS=sum(DailyEnrollAttend$Present)/sum(DailyEnrollAttend$Enrolled)*100, KAMS=NA,KCCP=NA, KBCP=NA)

AttRateYTDKAPSKAMS<-data.frame(WeekOfShortDateLabel="YTD KAPS & KAMS", KAPS=sum(subset(DailyEnrollAttend, SchoolInitials!="KCCP")$Present)/sum(subset(DailyEnrollAttend, SchoolInitials!="KCCP")$Enrolled)*100, KAMS=NA,KCCP=NA, KBCP=NA)
  
  
AttRateByWeekBySchool.table<-rbind(AttRateByWeekBySchool.table,AttRateYTDBySchool,AttRateYTDKAPSKAMS, AttRateYTDRegion)
AttRateByWeekBySchool.table[,c(2:5)] <- round(AttRateByWeekBySchool.table[,c(2:5)],1)  




names(AttRateByWeekBySchool.table)[1]<-"Week of" #a better column title
AttRateByWeekBySchool.xtable<-xtable(AttRateByWeekBySchool.table)

if(length(unique(DailyEnrollAttend$WeekOfShortDateLabel)) > 15){
  lastxweeks <- 15
  } else {lastxweeks <-length(unique(DailyEnrollAttend$WeekOfShortDateLabel))
  }
```
#### CPS Impact vs Powerschool YTD Attendance
```{r Impact_vs_PowerSchool_Att, results='asis'}
impact<-data.frame(School=c('KAPS',
                            'KAMS',
                            'KCCP',
                            'KBCP',
                            'KAPS/KAMS',
                            'Region'),
                   PowerSchool=c(95.9,
                                 95.6,
                                 96.5,
                                 96.6,
                                 95.8,
                                 95.9),
                   Impact=c(95.9,
                            95.9,
                            96.6,
                            96.7,
                            95.9,
                            96.1
                            ))
impact$Difference<-impact$Impact - impact$PowerSchool

kable(impact, format='html', table.attr='class="table table-responsive table-hover table-condensed table-striped"', output=TRUE)
```



#### Daily Enrollment and Attendance by School (last `r I(lastxweeks)` weeks)
The greeen line demarcates 95% of enrollment (i.e., our regional daily attendance goal).
```{r fig_Daily_Attendance, fig.height=3, fig.width = 7, dpi=150}
#out.height=225, out.width=525}

last15weeks<-ymd(as.character(floor_date(today() -weeks(20), unit="week")+1))

DailyEnrollAttend.plotdata<-subset(DailyEnrollAttend, Date>=last15weeks)
DailyEnrollAttend.plotdata$Day <- wday(DailyEnrollAttend.plotdata$Date)
DailyEnrollAttend.plotdata$Enrolled95Pct <- DailyEnrollAttend.plotdata$Enrolled*.95
DailyEnrollAttend.plotdata.melt<-melt(DailyEnrollAttend.plotdata, id=c("Date", "Day", "SchoolInitials", "WeekOfDateLabel"), measure.vars=c("Enrolled", "Enrolled95Pct", "Present"))

DailyEnrollAttend.plotdata.melt$variable<-factor(DailyEnrollAttend.plotdata.melt$variable, labels=c("Enrolled", "95% of Enrolled", "Attended"))


ggplot(DailyEnrollAttend.plotdata.melt, aes(x=Day, y=value)) + 
    geom_step(direction="hv", aes(color=variable)) + 
    scale_x_continuous(breaks = c(2,3,4,5,6), labels=c("M","T","W","R","F")) + #Change numberd week days to lettered
    scale_y_continuous("# of Students") + 
    scale_colour_manual("", values=c("#8D8685", "#439539", "black")) +
    facet_grid(SchoolInitials~WeekOfDateLabel, scales="free_y") +
    theme_bw() + 
    theme(legend.position="bottom", 
          strip.text.x=element_text(size=4.75),
          strip.text.y=element_text(size=4.75),
          axis.text.x=element_text(size=5),
          axis.text.y=element_text(size=4),
           axis.title.x=element_text(size=6),
          axis.title.y=element_text(size=6),
          legend.text=element_text(size=5)
          )


```

#### Weekly & YTD Attendance by School

```{r tbl_Weekly_Attendance, results='asis'}
#, dpi=300}
library(knitr)
kable(AttRateByWeekBySchool.xtable, format='html', table.attr='class="table table-responsive table-hover table-condensed table-striped"', output=TRUE)


```




#### Student Attendance 

##### KAPS
```{r Student_Attendance_KAPS, results='asis'}
AttByStudentBySchool<-arrange(ddply(Attendance, .(LASTFIRST, SchoolInitials, GRADE_LEVEL), summarise, Absences=sum(ABSENT)), SchoolInitials, desc(Absences))

AttByStudentBySchool<-na.omit(AttByStudentBySchool)
names(AttByStudentBySchool)<-c("Student", "School", "Grade", "Absences")

kable(AttByStudentBySchool[AttByStudentBySchool$School=="KAPS",][1:15,], format='html', table.attr='class="table table-responsive table-hover table-condensed table-striped"', output=TRUE)
```

##### KAMS
```{r Student_Attendance_KAMS, results='asis'}
kable(AttByStudentBySchool[AttByStudentBySchool$School=="KAMS",][1:15,], format='html', table.attr='class="table table-responsive table-hover table-condensed table-striped"', output=TRUE, row.names=FALSE)
```

##### KCCP
```{r Student_Attendance_KCCP, results='asis'}
kable(AttByStudentBySchool[AttByStudentBySchool$School=="KCCP",][1:15,], format='html', table.attr='class="table table-responsive table-hover table-condensed table-striped"', output=TRUE,  row.names=FALSE)
```

##### KBCP
```{r Student_Attendance_KBCP, results='asis'}
kable(AttByStudentBySchool[AttByStudentBySchool$School=="KBCP",][1:15,], format='html', table.attr='class="table table-responsive table-hover table-condensed table-striped"', output=TRUE,  row.names=FALSE)
```

### Suspensions
In this report the metric **suspensions** represents days absent due to suspensions and *not* the number of suspension events. That is, a student recieving a 3-day suspension would appear in this report as having 3 "suspensions" (i.e., three absenses due to being out of school serving a suspension).
```{r Suspensions, cache=FALSE, echo=FALSE, results='hide'}
Suspensions<-copy(data.table(subset(Attendance, ATT_CODE=='S')))

setnames(Suspensions, 
         c("GRADE_LEVEL", "CALENDARDATE", "LASTFIRST", "SchoolInitials" ),
         c("Grade", "Date", "Student", "School")   
         )

Suspensions[,WeekInYear:=lubridate::week(Suspensions$Date)]


Suspensions[,WeekInSchoolYear:=(floor_date(ymd_hms(Suspensions$Date), unit="week")-min(floor_date(ymd_hms(Suspensions$Date), unit="week")))/dweeks(1)+1]

Suspensions[,WeekOfDate:=floor_date(ymd_hms(Suspensions$Date), unit="week") + days(1)]

Suspensions<-copy(arrange(Suspensions, Date))

Suspensions[,WeekOfShortDateLabel:=paste(lubridate::month(Suspensions$WeekOfDate,label=TRUE, abbr=TRUE), day(Suspensions$WeekOfDate), sep=" ")]

Suspensions[,WeekOfShortDateLabel:=factor(Suspensions$WeekInSchoolYear, labels=unique(Suspensions$WeekOfShortDateLabel))]

Suspensions$School[Suspensions$SCHOOLID==7810]<-"KAMS"
Suspensions$School[Suspensions$SCHOOLID==78102]<-"KAPS"
Suspensions$School[Suspensions$SCHOOLID==400146]<-"KCCP"
Suspensions$School[Suspensions$SCHOOLID==400163]<-"KBCP"
Suspensions$School<-factor(Suspensions$School, levels=c("KAPS", "KAMS", "KCCP", "KBCP"))

DailySuspensionByGradeByWeek<-ddply(Suspensions, .(School, Grade, WeekOfShortDateLabel), summarise, Suspended=sum(ABSENT))


#weekly suspension by school 
WeeklySuspensionsBySchool.table<-cast(DailySuspensionByGradeByWeek, WeekOfShortDateLabel~School, sum, margins=TRUE)

# Change Week of row (all) to Total
levels(WeeklySuspensionsBySchool.table$WeekOfShortDateLabel)[levels(WeeklySuspensionsBySchool.table$WeekOfShortDateLabel)=="(all)"]<-"Total"

setnames(WeeklySuspensionsBySchool.table, c("WeekOfShortDateLabel","(all)"), 
                                          c("Week of", "Total"))
levels(WeeklySuspensionsBySchool.table[,1])[levels(WeeklySuspensionsBySchool.table[,1])=="(all)"]<-"Total"

WeeklySuspensionsBySchool.xtable<-xtable(WeeklySuspensionsBySchool.table, digits=0)


#YTD Suspsesnions by Grade By School
YTDSuspensionsByGradeBySchool.table<-cast(DailySuspensionByGradeByWeek, Grade~School, sum, margins=TRUE)

setnames(YTDSuspensionsByGradeBySchool.table, "(all)", "Total")

levels(YTDSuspensionsByGradeBySchool.table[,1])[levels(YTDSuspensionsByGradeBySchool.table[,1])=="(all)"]<-"Total"

YTDSuspensionsByGradeBySchool.xtable<-xtable(YTDSuspensionsByGradeBySchool.table, digits=0)



#YTD Suspsesnions by Grade by Week
YTDSuspensionsByWeekByGrade.table<-data.table(cast(DailySuspensionByGradeByWeek, WeekOfShortDateLabel~Grade, sum, margins=TRUE))

setnames(YTDSuspensionsByWeekByGrade.table, c("WeekOfShortDateLabel", "(all)"), c("Week of", "Total"))
levels(YTDSuspensionsByWeekByGrade.table[,1])[levels(YTDSuspensionsByWeekByGrade.table[,1])=="(all)"]<-"Total"

YTDSuspensionsByWeekByGrade.xtable<-xtable(YTDSuspensionsByWeekByGrade.table, digits=0)
```
#### Total YTD Suspensions 
##### by School 
```{r tbl_Susps_by_School, results='asis'}
kable(YTDSuspensionsByGradeBySchool.xtable, format='html', table.attr='class="table table-responsive table-hover table-condensed table-striped"', output=TRUE,  row.names=FALSE)
```
##### by Week

```{r tbl_Susps_by_week, results='asis'}
kable(YTDSuspensionsByWeekByGrade.xtable, format='html', table.attr='class="table table-responsive table-hover table-condensed table-striped"', output=TRUE,  row.names=FALSE)
```

#### YTD Students Suspended (ordered by number of suspensions)
```{r tbl_Suspensions_Leaderboard, results='asis'}
Sups.leaders<-arrange(ddply(Suspensions, .(Student, School, Grade), summarise, Suspensions=sum(ABSENT)), desc(Suspensions))
#names(Sups.leaders)<-c("Student", "School", "Grade", "Suspensions")


kable(Sups.leaders,format='html', table.attr='class="table table-responsive table-hover table-condensed table-striped"', output=TRUE,  row.names=FALSE)
```
