---
title: "Simulated Paths Proof of Concept"
author: "Chris Haid"
date: "October 1, 2014"
output: html_document
runtime: shiny
---

## [Create something today, even if it sucks!](http://themakeryproject.com/wp-content/uploads/2013/05/witandwhistlefathersday_mothersday_cards_2013_6.jpg)
### Or, how to quickly develop an application to show simulated versus acutal outcomes while highlighting where we'd really like your students to be.

#### How this works.

1. Pick a measurement Scale (i.e., a subject to lay people).
2. Pick a student.  Notice how nicely we've plotted their progress to day in that subject.  Neat.
3. Select at which grade you'd like simulated data to end.  10th is a good grade, I'd keep it there.
4. Check the __Show Simulated Data__, be patient,  and _have your mind blown!_
5. Check __Highligh highs__, be patient, again and *__have your mind blown, again!!__*
6. Fiddle with __Highlight simulated scores at or above:__ or the student or subject.
7. Check out the simulated data in the cool, sortable table below.
8. Now grab yourself a beer.  Think about the possibilities!

#### Visualization
```{r visualization, echo=FALSE}
require(mapvisuals)
require(lubridate)


#load fake MAP Data
data(nweamap)

gls<-function(x){
  out<-mapply(grade_level_season, x)
  out
}

# subest for brevity's sake
subjects<-c("Mathematics", "Reading")

map_results<-nweamap %>% 
  filter(MeasurementScale %in% subjects, 
         year(mdy(nweamap$TestStartDate)) >= 2010,
         Grade>=5
         ) %>% mapvizier



students<-map_results$mapData %>% 
  select(StudentID, StudentLastname, StudentFirstname, Grade) %>%
  group_by(StudentID, StudentLastname, StudentFirstname) %>%
  summarize(CurrentGrade=max(Grade)) %>% 
  mutate(StudentName=paste(StudentLastname, StudentFirstname, sep=", ")) %>%
  ungroup %>%
  select(StudentName, StudentID)

stus<-unlist(with(students, split(as.character(StudentID), StudentName)))



inputPanel(
  selectInput("ms", label = "Select Measurement Scale",
              choices = subjects, selected = "Reading"),
  
  selectInput("student", label = "Select Student:",
               choices = stus),
  
  selectInput("end_grade", label = "Select Simulation End Grade",
               choices = c(8:10), selected=10),
  
  checkboxInput("sim_yes", label = "Show Simulated Data",
                value=FALSE),
  
  br(),

      
  checkboxInput("hl_yes", label = "Highlight highs",
                value=FALSE),
  
  numericInput(inputId = "ontrack_rit", 
               value=250,
               label = "Highlight simulated scores at or above:",
               min = 195, max=300, step=10)

)


plot_data <- reactive({
  plot_data <- map_results$mapData %>% 
    filter(StudentID==as.integer(input$student), 
    MeasurementScale==input$ms) %>%
    mutate(GradeSeason=Grade + gls(Season))
  plot_data
  
})

start_grade <- reactive({
    min(plot_data()$Grade)
})

start_rit <- reactive({
    start_rit<-dplyr::filter(plot_data(), Grade==min(Grade), 
                             mdy(TestStartDate)==min(mdy(TestStartDate)))
    start_rit$TestRITScore
})

sim<-reactive({
  
  sim<-norm_sim(start.grade=start_grade(), 
                start.subject = input$ms,
                end.grade = input$end_grade,
                start.rit = start_rit()
                )
  
   sim<-sim %>% data.frame %>% mutate(Season=ifelse(Season==41,"Fall","Winter"),
              GradeSeason=StartGrade+gls(Season))
  
  sim_high <- filter(sim, GradeSeason==max(GradeSeason)) %>% 
    mutate(High=StartRIT>=input$ontrack_rit) %>%
    select(ID, High)
  
  sim <- sim %>% left_join(sim_high, by="ID")
  sim
})

renderPlot({
  
  p<-ggplot(plot_data(), aes(x=GradeSeason, y=TestRITScore))
  
  if(input$sim_yes){
    p<-  p + geom_line(data=sim(), 
              aes(x=GradeSeason, y=StartRIT, group=ID), 
              color='gray',
              alpha=.05) 
  } 
if(input$hl_yes){
  p <-  p + geom_line(data=filter(sim(), High==TRUE), 
              aes(x=GradeSeason, y=StartRIT, group=ID), 
              color='lightgreen',
              alpha=.3) 
}  
p <- p +
  geom_line(color='blue') +
  theme_bw() + 
  ylab("RIT Score") +
  xlab("Grade") 
  
p
    
})

```
#### Data Table
```{r datatable, echo=FALSE}
renderDataTable({
  sim()
  })

```




