---
title: "2018-02-07 PMM Culture"
output: html_notebook
---

## Package set-up

We are pulling and visualizing DeansList Data, so we need `silounloadr` and `tidyverse` packages
```{r set-up}
library(silounloadr)
library(tidyverse)
library(lubridate)
library(kippcolors)
```

## Getting DL Data

This should be a straightforward pull from BW:

```{r get_dl_data}
dl_incidents <-get_deanslist(table_name = "incidents")
dl_behaviors <-get_deanslist("behaviors")

glimpse(dl_behaviors)

```

We'll need PS data to match students to grade and homeroom
```{r}
ps_students <- get_powerschool("students")

glimpse(ps_students)

students <- ps_students %>%
  select(studentid = student_number, home_phone, grade_level, schoolid) %>%
  collect()

```



Let's focus on behaviors for now, since we don't normally for PMM.
```{r}
behaviors<-dl_behaviors %>%  
  select(studentid = studentschoolid,
         schoolname,
         behaviordate, 
         behavior,
         behaviorcategory,
         pointvalue,
         stafffirstname,
         stafflastname, 
         ) %>%
  filter(behaviordate >= '2017-08-15 00:00') %>%
  collect()

```
```{r}

behaviors_stus <- students %>%
  inner_join(behaviors, by = 'studentid')

behav_summary <- behaviors_stus %>%
  mutate(weekdate = floor_date(behaviordate, unit = 'week')) %>%
  filter(behaviorcategory != "Character Strengths") %>%
  group_by(schoolname, grade_level, weekdate) %>%
  summarize(count = n(),
            total_deductions = -1*sum(pointvalue),
            avg_deductions = mean(-1*pointvalue)) %>%
  filter(count >= 5) %>%
  mutate(school = abbreviate(schoolname, minlength = 3L)
         )
```


```{r}

symbol_size = 1
behav_summary %>%
  filter(grade_level > 4) %>%
ggplot(aes(x=weekdate, y = total_deductions)) +
  geom_point(aes(color = total_deductions), size = symbol_size) +
  geom_segment(aes(xend = weekdate, yend=0, color = total_deductions), size = symbol_size) +
  facet_wrap(~ school  + grade_level) +
  theme_kipp_light() +
  scale_color_kipp(discrete = FALSE) +
  labs(x = "Week",
       y = "Total Deductions",
       color = "Total Deductions")
```

```{r}
symbol_size = 1
behav_summary %>%
  filter(grade_level > 4) %>%
ggplot(aes(x=weekdate, y = count)) +
  geom_point(aes(color = count), size = symbol_size) +
  geom_segment(aes(xend = weekdate, yend=0, color = count), size = symbol_size) +
  facet_wrap(~ school  + grade_level) +
  theme_kipp_light() +
  scale_color_kipp(discrete = FALSE) +
  labs(x = "Week",
       y = "Count of Deductions",
       color = "Count of Deductions")
```

```{r}
symbol_size = 1
behav_summary %>%
  filter(grade_level > 4) %>%
ggplot(aes(x=weekdate, y = avg_deductions)) +
  geom_point(aes(color = avg_deductions), size = symbol_size) +
  geom_segment(aes(xend = weekdate, yend=0, color = avg_deductions), size = symbol_size) +
  facet_wrap(~ school  + grade_level) +
  theme_kipp_light() +
  scale_color_kipp(discrete = FALSE) +
  labs(x = "Week",
       y = "Avg Deduction",
       color = "Avg Deduction")
```

# Transfers

Let's start rethinking visuals and tables for this (pulling this code from IDEA)
```{r transfer_helper_functions}

force_weekday <- function(date, offset = 0) {

  date <- lubridate::ymd(date)

  if(lubridate::wday(date) == 1) date <- date + lubridate::days(1 + offset) # Sunday
  if(lubridate::wday(date) == 7)  date <- date + lubridate::days(2 + offset) # Saturday

  # return
  date

}

# Get membership on a given date

get_membership_on_date <- function(con, date, offset = 0) {

  # Ensure date is a weekday and then format
  date <- force_weekday(date, offset)

  #format date to YYYY-MM-DD
  date <- lubridate::parse_date_time(date, "ymd")
  date_formatted <- format(date, "%Y-%m-%d")

  qry <- sprintf(
          "SELECT * FROM membership WHERE CALENDARDATE = '%s'",
          date_formatted
         )

  # return
  tbl(con, sql(qry))





}



# add schools' names
school_names <- function(school_number, as_factor = TRUE){

  names_lookup <- function(x) {
    switch(as.character(x),
           "78102" = "KAP",
           "7810" = "KAMS",
           "400146" = "KCCP",
           "400163" = "KBCP",
           "4001802" = "KOP",
           "400180" = "KOA")
  }
  out <- vapply(school_number, FUN = names_lookup, "character")
  if(as_factor) out <- factor(out, levels = c("KAP", "KAMS", "KCCP", "KBCP", "KOP", "KOA"))

  out

}

```

```{r get_transfers}




students <-get_ps("students")

hsr_dates <- c(ymd("171002"), ymd("161004"), ymd("151001"))


membs_list <- hsr_dates %>%
  map(~get_membership_on_date(silo_dbname_ps_mirror, .) %>%
  collect())

enrolled <- bind_rows(membs_list)

names(enrolled) <- tolower(names(enrolled))

students <- collect(students)

names(students) <- tolower(names(students))
```

```{r transfer_processing}
enrolled_2 <- enrolled %>%
  left_join(students, by = c("studentid"="id")) %>%
  mutate(
    date_start =  ymd_hms(calendardate),
    date_end = date_start + years(1),
    sy = sprintf("%s-%s",
                 year(date_start),
                 year(date_end)),
    transferred = (!is.na(exitcode) &
                     exitcode != "GR" &
                     ymd_hms(exitdate) < date_end),
    exit_date = ymd_hms(ifelse(ymd_hms(exitdate) >= date_end, NA, exitdate)),
    exit_month = month(ymd_hms(exitdate),label = TRUE, abbr = TRUE)
    )


transfers_by_day <- enrolled_2 %>%
  group_by(sy, schoolid.x,  exit_date) %>%
  summarize(transfers = sum(transferred)) %>%
  group_by(sy, schoolid.x) %>%
  mutate(cum_transfers = dplyr::order_by(
    exit_date,
    cumsum(as.integer(transfers))))

transfers_by_day_by_code <- enrolled_2 %>%
  filter(transferred) %>%
  group_by(sy, schoolid.x,  exit_date, exitcode) %>%
  summarize(transfers = sum(transferred)) %>%
  group_by(sy, schoolid.x, exitcode) %>%
  mutate(cum_transfers = dplyr::order_by(
    exit_date,
    cumsum(as.integer(transfers))))

transfers_by_month <- transfers_by_day %>%
  mutate(exit_month = month(exit_date, label = TRUE, abbr = TRUE)) %>%
  group_by(sy, schoolid.x, exit_month) %>%
  summarize(
    date = max(exit_date),
    transfers = sum(transfers),
    cum_transfers = max(cum_transfers))

transfers_by_month_by_code <- transfers_by_day_by_code %>%
  mutate(exit_month = month(exit_date, label = TRUE, abbr = TRUE)) %>%
  group_by(sy, schoolid.x, exit_month, exitcode) %>%
  summarize(
    date = max(exit_date),
    transfers = sum(transfers),
    cum_transfers = max(cum_transfers))


transfer_order <- c("Dropped Out", "Expelled", "Behavior/Discipline", "Academics", "Avoid Retention", "Special Needs", "Other", "Don't Know", "Xfer Other KIPP", "Transport", "Moved")

transfer_reasons<-data_frame(exitcode=as.character(c(1:11)),reason=c("Dropped Out", "Moved", "Transport", "Expelled", "Behavior/Discipline", "Academics", "Avoid Retention", "Special Needs", "Other", "Don't Know", "Xfer Other KIPP"))

transfer_reasons <- transfer_reasons %>%
  mutate(reason = factor(reason, levels = transfer_order, ordered = TRUE))

month_order <- c("Oct", "Nov", "Dec", "Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep")

month_factor <- factor(month_order, levels = month_order, ordered = TRUE)

scaffold <- expand.grid(sy = unique(transfers_by_day_by_code$sy),
                        schoolid.x = unique(transfers_by_day_by_code$schoolid.x),
                        exit_month = month_factor,
                        exitcode = unique(transfers_by_day_by_code$exitcode),
                        stringsAsFactors = FALSE)


transfers_by_month_2 <- scaffold %>%
  left_join(transfers_by_month_by_code, by =c("sy", "schoolid.x", "exit_month", "exitcode")) %>%
  left_join(transfer_reasons, by="exitcode")%>%
  group_by(sy, schoolid.x,  exitcode) %>%
  mutate(cum_transfers_2 = as.integer(zoo::na.locf(cum_transfers, na.rm = FALSE)),
         month = factor(exit_month, levels = month_order, ordered=  TRUE)) %>%
  ungroup() %>%
  mutate(sy = factor(sy, levels = rev(unique(sy)), ordered = TRUE),
         school_name = school_names(schoolid.x)
         ) %>%
  arrange(sy, school_name, month, desc(reason))



transfer_goals <- enrolled %>%
  group_by(schoolid, calendardate) %>%
  summarize(N = n()) %>%
  mutate(yearly_goal = round(.1 * N),
         monthly_goal = yearly_goal/12,
         sy2 = sprintf("%s-%s",
                      year(ymd_hms(calendardate)),
                      year(ymd_hms(calendardate))+1),
         sy = factor(sy2, levels=rev(unique(sy2)), ordered = TRUE),
         schoolid.x=schoolid,
         school_name = school_names(schoolid)
  )
```


```{r}
View(transfers_by_month_2)

transfer_goals 

```

