---
title: "Spring 2018-2018 MAP Trackers"
output: html_notebook
---

Load packages 
```{r packages}
library(tidyverse)
library(silounloadr)

library(mapvizieR)
library(purrr)
library(xlsx)
library(janitor)
```

Paramters adn functions are set here
```{r paramaters}
todays_date<-lubridate::today()

sy <- calc_academic_year(todays_date, format = "second_year")
sy_short <- stringr::str_extract(sy, "\\d{2}$")
#sy <- 2018
#school <- 'KCCP'
suppress_kinder <- FALSE
subjs <- c("Reading", "Mathematics")
```

And local functions are here:

```{r local_functions}
# Function to separte combined cdf to roster and results
separate_cdf <- function(combinded_cdf, district_name = "Not provided"){
  ar_names <- names(ex_CombinedAssessmentResults) %>% tolower()
  stu_names <- names(ex_CombinedStudentsBySchool) %>% tolower()

  if (!"districtname" %in% tolower(names(combinded_cdf))) {
    combinded_cdf <- combinded_cdf %>% mutate_(districtname = ~district_name)
  }

  roster<-combinded_cdf %>%
    select_(.dots = stu_names) %>%
    unique

  cdf<-combinded_cdf %>% select(-studentlastname:-studentfirstname,
                                -studentmi:-studentgender,
                                -grade) %>%
    mutate(testid=as.character(testid))

  out <- list(cdf = cdf,
              roster = roster)

}
```

And now we pull down our data

```{r}
nwea <- get_nwea_map("cdf_combined_kipp_cps")


#pull_terms <- list("Spring 2016-2017", "Fall 2017-2018", "Winter 2017-2018")

pull_terms <- list("Spring 2017-2018", "Fall 2018-2019", "Winter 2018-2019")

nwea_2 <-pull_terms %>%
  map_df(~ {nwea %>% 
              filter(term_name == .x) %>%
              collect(allowLargeResults = TRUE)
            }
         )

old_names <- names(nwea_2)
new_names <- old_names %>% stringr::str_replace_all("_", "")
names(nwea_2) <- new_names

nwea<-nwea_2 %>% select(-ClassName, -TeacherName) %>% 
   mutate(teststartdate = lubridate::mdy(teststartdate), 
          teststarttime = lubridate::hms(teststarttime)) %>%
  mutate(teststarttime = as.character(teststarttime) %>% stringr::str_replace_all('[HM]\\W', ":") %>% stringr::str_replace("S", ""),
         teststartdate = as.character(teststartdate)) %>% 
  distinct()

  

nwea <- nwea %>%
  #filter(grepl("Create", SchoolName)) %>%
  mutate(grade=as.integer(ifelse(grade=="K", 0, grade)),
         measurementscale=ifelse(measurementscale=="Science - General Science",
                                 "General Science",
                                 measurementscale)
  )

# get students

current_ps_roster <- get_powerschool("students")

current_ps_roster <- current_ps_roster %>%
  filter(enroll_status == 0) %>%
  collect()

current_ps_roster <- current_ps_roster %>%
  clean_names() %>%
  select(studentid = student_number,
         studentlastname = last_name,
         studentfirstname = first_name,
         schoolid,
         grade_level,
         home_room)

current_stus_non_k<-current_ps_roster %>% filter(grade_level != 0) %>%
  select(studentid, current_grade = grade_level)

# Get spring


last_spring <- sprintf("Spring %s-%s", sy-2, sy-1)
this_fall <- sprintf("Fall %s-%s", sy-1, sy)
this_winter <- sprintf("Winter %s-%s", sy-1, sy)

map_spring <- nwea %>%
  filter(termname==last_spring,
         measurementscale %in% subjs) %>%
  inner_join(current_stus_non_k, by = "studentid") %>%
  mutate(
    Orignal_Score=testritscore,
    Original_Percentile=testpercentile,
    Original_Term = sprintf("Spring %s", sy-1),
    Fall_as_Spring = FALSE)


# find current students with missing spring scores ###


current_stus_no_spring <- map_spring %>%
  split(.$measurementscale) %>%
  map(~anti_join(current_ps_roster,
                 .,
                 by=c("studentid")
  ) %>%
    mutate(measurementscale = unique(.x$measurementscale)) %>%
    select(studentid, measurementscale) %>%
    unique) %>%
  bind_rows

# Get fall
map_fall <- nwea %>%
  filter(termname==this_fall,
         grade != 0,
         measurementscale %in% subjs) %>%
  inner_join(current_stus_no_spring,
             by = c("studentid", "measurementscale")) %>%
  mutate(Current_Grade=grade,
         Orignal_Score=testritscore,
         Original_Percentile=testpercentile,
         Original_Term = sprintf("Fall %s", sy-1),
         Fall_as_Spring = TRUE,
         termname = last_spring)

# get fall kinder
 current_stus_kinder<-current_ps_roster %>%
   filter(grade_level == 0) %>%
   select(studentid)
 
 map_fall_kinder <- nwea %>%
   filter(termname==this_fall,
          measurementscale %in% subjs,
         studentid %in% current_stus_kinder$studentid
  ) %>%
  mutate(Current_Grade=grade,
         Orignal_Score=testritscore,
         Original_Percentile=testpercentile,
         Original_Term = sprintf("Fall %s", sy-1),
         Fall_as_Spring = NA,
         CPS_Imputed = NA)



# Check if spring is na.  If so updated Fall_as_Spring indicator.
#If Fall_as_Spring==T then update TestRITScore to Fall Score
# NEW if KAP 5 Then updated Fall_as_Spring to FALSE, update
# Imputed_from_Fall to TRUE and filter those students, ipmute for
# math and reading, bind, then join

map_fall_reading <-  map_fall %>%
  filter(measurementscale == "Reading") %>%
  mutate(testritscore = as.integer(sqrpr::cps_equate(testritscore, measurementscale, grade)),
         CPS_Imputed = TRUE,
         grade = grade -1)

map_fall_math <-  map_fall %>%
  filter(measurementscale == "Mathematics") %>%
  mutate(testritscore = as.integer(sqrpr::cps_equate(testritscore, measurementscale, grade)),
         CPS_Imputed = TRUE,
         grade = grade -1)


map_joined <- dplyr::bind_rows(map_spring %>%
                                  mutate(CPS_Imputed = FALSE),
                                map_fall_reading,
                                map_fall_math
                               ) %>%
  group_by(studentid, termname) %>%
  filter(grade == min(grade)) %>% #need to remove retained students
  ungroup
```


Calculating stuff
```{r}

cdf <- separate_cdf(map_joined, "KIPP Chicago")

# add missing columns 
cdf$cdf <- cdf$cdf %>% 
  mutate(goal7name = NA, 
         goal7ritscore = NA, 
         goal7stderr = NA, 
         goal7range = NA, 
         goal7adjective = NA, 
         goal8name = NA, 
         goal8ritscore = NA, 
         goal8stderr = NA, 
         goal8range = NA, 
         goal8adjective = NA, 
         projectedproficiencystudy3 = NA, 
         projectedproficiencylevel3 = NA)

map_mv<-mapvizieR(cdf = cdf$cdf, roster = cdf$roster)

cdf_k <- separate_cdf(map_fall_kinder %>%
                        mutate(teststarttime = as.character(teststarttime),
                               goal7name = NA,
                               goal7ritscore = NA,
                               goal7stderr = NA,
                               goal7range = NA,
                               goal7adjective = NA,
                               goal8name = NA,
                               goal8ritscore = NA,
                               goal8stderr = NA,
                               goal8range = NA,
                               goal8adjective = NA,
                               projectedproficiencystudy3 = NA,
                               projectedproficiencylevel3 = NA), "KIPP Chicago")

map_mv_k<-mapvizieR(cdf = cdf_k$cdf, roster = cdf_k$roster)


# reudce growth to just "start_only terms"
growth_df <- map_mv$growth_df %>% 
#dplyr::ungroup() %>%
#as.data.frame() %>%
  filter(match_status == "only start") %>%
  inner_join(map_mv$cdf %>% select(testid,
                                   current_grade:cps_imputed,
                                   testedatkipp),
             by=c("start_testid" = "testid")) %>%
  inner_join(map_mv$roster, by = c("studentid"))


growth_df_k <- map_mv_k$growth_df %>%
  filter(match_status == "only start",
         growth_window == "Fall to Spring") %>%
  inner_join(map_mv_k$cdf %>% select(testid,
                                     current_grade:cps_imputed,
                                     testedatkipp),
             by=c("start_testid" = "testid")) %>%
  inner_join(map_mv_k$roster, by = c("studentid"))


  growth_df_joined <- current_ps_roster %>%
   left_join(bind_rows(growth_df, growth_df_k),
             by = c("studentid" = "studentid"))

growth_df_joined <- current_ps_roster %>%
  select(studentid) %>%
 left_join(growth_df_joined,
            by = c("studentid"))


growth_df_joined %>% names()
```


Update school names
```{r school_names}
school_names <- data.frame(schoolid=c(78102, 7810, 400146, 4001632, 400163, 4001802, 400180),
                           School = c("KIPP Ascend Primary School",
                                          "KIPP Ascend Middle School",
                                          "KIPP Academy Chicago",
                                          "KIPP Bloom Primary",
                                          "KIPP Bloom College Prep",
                                          "KIPP One Primary",
                                          "KIPP One Academy"
                                          )
)


growth_df_joined <- growth_df_joined %>% left_join(school_names, by = "schoolid")

# growth_df_joined$schoolname<-
#   school_names$schoolname[match(growth_df_joined$schoolid,
#                                 school_names$schoolid)]

```

Now we prep the mapvizier file to take get ready for the excel maniuplation (i.e., getting kids with goals and results all lined up)

```{r}
map_growth_final_1<-growth_df_joined %>%
  mutate(
    #Goal = ifelse(Current_Grade==0, R42, R22),
    TypicalGrowth_Spring=start_testritscore + reported_growth,
    TieredGrowth_Spring= start_testritscore + accel_growth,
    TypicalGrowth_Winter=start_testritscore + round(reported_growth/2),
    TieredGrowth_Winter=start_testritscore + round(accel_growth/2),
    SchoolInitials = abbrev(School, exceptions = list(old ="KAPS", new="KAP")),
    #Fall_Score = ifelse(original_term == "Fall 2015", orignal_score, NA),
    Fall_Score = orignal_score,
    Fall_as_Spring = fall_as_spring) %>%
  select(SchoolInitials,
         MeasurementScale = measurementscale.x,
         Grade=grade_level,
         StudentID=studentid,
         StudentFirstName = studentfirstname.x,
         StudentLastName = studentlastname.x,
         Spring_Score = start_testritscore,
         Fall_Score,
         TypicalGrowth_Winter,
         TieredGrowth_Winter,
         TypicalGrowth_Spring,
         TieredGrowth_Spring,
         Fall_as_Spring,
         Tested_at_KIPP = testedatkipp,
         CPS_imputed = cps_imputed
         #,Spring_Imputed

  ) %>%
  mutate(Spring_Score = ifelse(Grade==0, NA, Spring_Score)) %>% # fix for kinder
  arrange(SchoolInitials, MeasurementScale, Grade, StudentLastName, StudentFirstName)


# get original fall scores

map_fall_actual <- nwea %>%
  filter(termname==this_fall,
         measurementscale %in% subjs) %>%
  select(StudentID = studentid,
         MeasurementScale = measurementscale,
         Fall_RIT = testritscore)

# get original winter scores
map_winter_actual <- nwea %>%
  filter(termname==this_winter,
         measurementscale %in% subjs) %>%
  select(StudentID = studentid,
         MeasurementScale = measurementscale,
         Winter_Score = testritscore)

map_growth_final<-map_growth_final_1 %>%
  left_join(map_fall_actual %>%
              select(StudentID,
                     MeasurementScale,
                     Fall_RIT),
            by = c("StudentID", "MeasurementScale")) %>%
  mutate(Fall_Score = Fall_RIT) %>%
  select(-Fall_RIT) %>%
  left_join(map_winter_actual %>%
              select(StudentID,
                     MeasurementScale,
                     Winter_Score),
            by = c("StudentID", "MeasurementScale")) %>%
  select(-StudentID) %>%
  arrange(SchoolInitials, MeasurementScale, Grade, StudentLastName, StudentFirstName) %>%
  unique()


todays_date<-format(lubridate::today(), "%y%m%d")
readr::write_csv(map_growth_final, path=paste0("../reports/Tracker_",todays_date,".csv"), na="")
```

## Excel Manipulation

This is where the magic happens.  The excel templates must have tabs for each grade subject. 

```{r fill_in_excel_templates}
schools <- map_growth_final$SchoolInitials %>% unique
#schools <- "KBCP"
#schools <- "KOA"
for(school in schools) {

  file_in <- sprintf("../Trackers/templates/MAP Tracker - SY%s %s.xlsx", sy_short, school)
  xls_wb<-loadWorkbook(file=file_in)
  sheets<-getSheets(xls_wb)

  map_school<-
    map_growth_final %>%
    filter(SchoolInitials==school,
           !is.na(MeasurementScale))

  for (subj in unique(map_school$MeasurementScale)){
    map_subj<-map_school %>% filter(MeasurementScale==subj)
    for(gr in unique(map_subj$Grade)){

      map_out <- map_subj %>%
        filter(Grade==gr)



      subj<-ifelse(subj=="General Science", "Science", subj)
      sheet_name <- paste(school, gr, subj)

      message(paste("Getting sheet:", sheet_name))
      sheet<-sheets[[sheet_name]]

      message("Adding name, spring and fall scores and winter goals  and scores")
      addDataFrame(map_out %>%
                     dplyr::select(StudentFirstName,
                            StudentLastName,
                            Spring_Score,
                            Fall_Score,
                            TypicalGrowth_Winter,
                            TieredGrowth_Winter,
                            Winter_Score
                     ) %>%
                     as.data.frame(),
                   sheet=sheet,
                   col.names=FALSE,
                   row.names=FALSE,
                   startRow=4,
                   startColumn=1
      )


      message("Adding spring goals")
      addDataFrame(map_out %>%
                     select(
                       TypicalGrowth_Spring,
                       TieredGrowth_Spring
                     ) %>%
                     as.data.frame(),
                   sheet=sheet,
                   col.names=FALSE,
                   row.names=FALSE,
                   startRow=4,
                   startColumn=10
      )


      message("Adding score source data")
      addDataFrame(map_out %>%
                     select(Fall_as_Spring,
                            Tested_at_KIPP,
                            CPS_imputed
                     ) %>%
                     as.data.frame(),
                   sheet=sheet,
                   col.names=FALSE,
                   row.names=FALSE,
                   startRow=4,
                   startColumn=15
      )



    }
  }
  message("Saving file.")
  file_out <- sprintf("../Trackers/complete/MAP Tracker - SY%s %s %s.xlsx", sy_short, school, todays_date)
  saveWorkbook(xls_wb, file=file_out)

}
```

