ANet Interim 3 Analysis
========================================================
```{r options, echo=FALSE}
options(replace.assign=TRUE,width=60)
opts_chunk$set( 
               fig.align='left', 
               #dev='tikz', 
               fig.width=11, 
               fig.height=4.5, 
               fig.show='hold', 
               cache=FALSE, 
               par=TRUE,
               echo=FALSE,
               message=FALSE,
               warning=FALSE,
               autodepend=TRUE
               )
```


```{r load_libraries, echo=FALSE}
setwd("~/Dropbox/Consulting/KIPP Ascend/Data Analysis/ANet/")

library(plyr)  #To manipulate data
library(reshape) #More data manipulation
library(ggplot2) #Graphics of grammer graphing
library(grid) #More Graphing
library(gridExtra) #better than par(mfrow=c(r,c)) for arranging ggplot2 and lattice graphics
library(lubridate) #for dealing with dates
library(xtable) #for pretty tables
```


```{r prep_ANet_Data, echo=FALSE}

#Get data from CSV file compiled from ANet reporting website
ANet.A3<-read.csv('/Users/chaid/Dropbox/Consulting/KIPP Ascend/Data Analysis/ANet/Results/ANET_Results_A1A2_130225.csv')
kippcols=c("#439539", "#BCD631", "#CFCCC1")


ANet.A3$School<-factor(ANet.A3$School, levels=c("KAP/MS", "KCCP", "Anet"), labels=c("KAP/MS", "KCCP", "ANet"))
ANet.A3$Subject<-factor(ANet.A3$Subject, levels=c("Math", "ELA"))
ANet.A3$Grade<-factor(ANet.A3$Grade, levels=c(2,5,6,7,8))
ANet.A3$Test<-factor(ANet.A3$Test, levels=c("A1","A2","A3", "A1A2", "A1A2A3"), labels=c("A1", "A2", "A3", "A1 & A2 Combined", "A1-A3 Combined"))
```

## Interim A1 - A3. Percentage points above or below the ANet networks average.

This chart simply shows the difference in our average interim score versus the network average.  Positive (green) values indicate outperforming the network average; negative (orange) values indicate underperformance. These data are arranged by interim.
```{r prep_overunder, echo=FALSE, warning=FALSE}
#Create plot showing performance above or below average

ANet.A3.melt<-melt(ANet.A3)

ANet.Percent.cast<-cast(ANet.A3.melt, Test + Grade + Subject ~ School + variable, subset=variable=="Percent")
names(ANet.Percent.cast)<-c("Test", "Grade", "Subject", "Ascend", "Create", "ANet")

ANet.Percent<-within(ANet.Percent.cast, {
  KAPMS<-(Ascend - ANet)
  KCCP <- (Create-ANet)
  variable<-"Percent" 
       })

ANet.Sixty.cast<-cast(ANet.A3.melt, Test + Grade + Subject ~ School + variable, subset=variable=="Pct60Above")
names(ANet.Sixty.cast)<-c("Test", "Grade", "Subject", "Ascend", "Create", "ANet")

ANet.Sixty<-within(ANet.Sixty.cast, {
  KAPMS<-Ascend - ANet 
  KCCP <- Create-ANet
  variable<-"Pct60Above" 
       })

ANet.Diffs<-rbind(ANet.Percent, ANet.Sixty)

ANet.Diffs<-ANet.Diffs[,-c(4:6)]

ANet.Diffs.melt<-melt(as.data.frame(ANet.Diffs), measure=c("KCCP", "KAPMS"), variable_name="School", na.rm=TRUE)#use as.data.frame() trick to strip deleted melt info per hadely's suggestion at https://stat.ethz.ch/pipermail/r-help/2009-January/185755.html

ANet.Diffs.melt$Pos<-ANet.Diffs.melt$value>0


ANet.Diffs.melt$School<-factor(ANet.Diffs.melt$School, levels=c("KAPMS", "KCCP"), labels=c("KAP/MS", "KCCP"))
ANet.Diffs.melt$Subject<-factor(ANet.Diffs.melt$Subject, levels=c("Math", "ELA"))
ANet.Diffs.melt$Grade<-factor(ANet.Diffs.melt$Grade, levels=c(2,5,6,7,8))
ANet.Diffs.melt$Test<-factor(ANet.Diffs.melt$Test, levels=c("A1","A2","A3","A1A2", "A1A2A3"), labels=c("A1", "A2", "A3", "A1 & A2 Combined", "A1-A3 Combined"))
```

```{r plot_overunder, fig.height=4.5, fig.width=10, echo=FALSE, warning=FALSE}

Anet.sub<-subset(ANet.Diffs.melt,variable=="Percent" & School!="KCCP" & Test!="A1 & A2 Combined")

ggplot(Anet.sub, aes(x=as.factor(Grade), y=value)) +
  geom_bar(aes(fill=Pos), position="dodge", stat="identity" ) + 
  geom_text(data = subset(Anet.sub, value>=0), aes(y=value+2, label=round(value), color=Pos), vjust=0) +
    geom_text(data = subset(Anet.sub, value<0), aes(y=value-2, label=round(value), color=Pos), vjust=1) +
  geom_hline(yintercept=0, show_guide=T) +
  facet_grid(Subject ~ Test) + 
  scale_fill_manual(values=c("#E27425","#439539"), guide="none") +  
  scale_colour_manual(values=c("#E27425","#439539"), guide="none") +
  labs(x="Grade", y="Points Above/Below Network Average") +
  theme_bw() + 
  expand_limits(y=c(-10,25)) +   
  ggtitle("ANet Interim 3: Percentage Points Above/Below Network Average")
```
This next polot shows the same difference in our average interim score versus the network average. Howver, I've grouped the data by grade level, so we can see performace by grade relative to the network over time. Again, positive (green) values indicate outperforming the network average; negative (orange) values indicate underperformance.
```{r plot_overunder_by_grade}
ggplot(subset(ANet.Diffs.melt,variable=="Percent" & School!="KCCP" & Test!="A1 & A2 Combined" & Test!="A1-A3 Combined"), aes(x=as.factor(Test), y=value)) +    geom_text(data = subset(ANet.Diffs.melt,variable=="Percent" & School!="KCCP" & Test!="A1 & A2 Combined" & Test!="A1-A3 Combined" & value>=0), aes(y=value+2, label=round(value), color=Pos), vjust=0) + geom_line(aes(x=as.numeric(Test))) + 
    geom_point(aes(color=Pos), size=5) + 
    geom_text(data = subset(ANet.Diffs.melt,variable=="Percent" & School!="KCCP" & Test!="A1 & A2 Combined" & Test!="A1-A3 Combined" & value<0), aes(y=value-2, label=round(value), color=Pos), vjust=1) +
    geom_hline(yintercept=0, show_guide=T) +
    facet_grid(Subject ~ Grade) + 
    scale_fill_manual(values=c("#E27425","#439539"), guide="none") +  
    scale_colour_manual(values=c("#E27425","#439539"), guide="none") +
    labs(x="Interim Assessment", y="Points Above/Below Network Average") +
    theme_bw() + 
    expand_limits(y=c(-10,25)) +   
    ggtitle("ANet Interim 3: Percentage Points Above/Below Network Average")
```

## Interim 1-3 Combined.  Average Score by grade and subject.
```{r plot_Anet3_Combined, fig.height=4.5, fig.width=10, echo=FALSE, warning=FALSE}

ANet.A3.combined.plot<-ggplot(subset(ANet.A3, Test=="A1-A3 Combined"), aes(x=Grade, y=Percent)) + 
  geom_text(aes(y=Percent+1,label=Percent, color=School), position = position_dodge(width=1), vjust=0, size=6) +
  geom_bar(aes(fill=School), position="dodge", stat='identity') + 
  facet_grid(Subject ~ .) +
  scale_fill_manual("",values=kippcols) +
  scale_colour_manual("",values=kippcols) +
  ylim(0,80) + 
  theme_bw() + 
  theme(legend.position="bottom") +
  ggtitle("ANet Interims 1,2 & 3 - Combined Results")
  
ANet.A3.combined.plot 
```
This shows the average of student scores, which themselves are the percent of question answered correctly.
```{r plot_Anet1_2_3, echo=FALSE}
ts<-5

ANet.A3.plot<-ggplot(subset(ANet.A3, Test!="A1 & A2 Combined" & School!="KCCP"), aes(x=Grade, y=Percent)) + 
  geom_text(aes(y=Percent+1,label=Percent, color=School), position = position_dodge(width=1), vjust=0, size=ts) +
  geom_bar(aes(fill=School), position="dodge", stat='identity') + 
  facet_grid(Subject ~ Test) +
  scale_fill_manual("",values=kippcols) +
  scale_colour_manual("",values=kippcols) +
  ylim(0,85) + 
  theme_bw() + 
  theme(legend.position="bottom") + 
  ggtitle("ANet Interims  - Percent of Questions Correct")
  
ANet.A3.plot 
```
This chart shows the *percent of students* earning a 60% or better.  
```{r plot_Anet1_2_3_60plus, echo=FALSE}
ANet.A3.pct60.plot<-ggplot(subset(ANet.A3, Test!="A1 & A2 Combined" & School!="KCCP"), aes(x=Grade, y=Pct60Above)) + 
  geom_text(aes(y=Pct60Above+1,label=Pct60Above, color=School), position = position_dodge(width=1), vjust=0, size=ts) +
  geom_bar(aes(fill=School), position="dodge", stat='identity') + 
  facet_grid(Subject ~ Test) +
  scale_fill_manual("",values=kippcols) +
  scale_colour_manual("",values=kippcols) +
  ylim(0,105) + 
  theme_bw() + 
  theme(legend.position="bottom") +
  ylab("Percent > 60%") +
  ggtitle("ANet Interims  - Percent of Students w/ Scores > 60 Percent")
  
ANet.A3.pct60.plot 

```